---
title: "Microbiome in Liver Transplant Manuscript"
author: "Nicholas P. Dylla"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: yeti
    highlight: pygments
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
editor_options: 
  chunk_output_type: console
header-includes:
  - |
    ```{=latex}
    \usepackage{fvextra}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{
      showspaces = false,
      showtabs = false,
      breaklines,
      commandchars=\\\{\}
    }
    ```
---

# {.tabset}
## Package Loading and Data Importing
### Load Packages and Custom Functions
```{r, package-loading, message=FALSE, warning=FALSE}
# BiocManager::install("mixOmicsTeam/mixOmics")
# remotes::install_github("ying14/yingtools2")
# remotes::install_github("mikabr/ggpirate")
# devtools::install_github("EmilHvitfeldt/paletteer")

library(pacman)
p_load(
  factoextra,               # Dimension reduction
  FactoMineR,               # Dimension reduction
  stabs,                    # Stability variable selection
  mboost,                   # Gradient boosting for model building
  ggrepel,                  # Visualization, repels labels on plots
  bestglm,                  # Logistic regression
  knitr,                    # To change R markdown PDF options
  caret,                    # To calculate model paramters
  cutpointr,                # Calculate cutpoints
  RPostgreSQL,              # Connect to PostgreSQL server
  phyloseq,                 # Phyloseq data wrangling
  microbiomeMarker,         # LEfSe analysis
  ComplexHeatmap,           # Heatmap generator
  paletteer,                # Extensive color palette
  tidyverse,                # Data wrangling and visualization
  circlize,                 # Color ramp builder
  rstatix,                  # Tidyverse statistics
  EnhancedVolcano,          # Volcano plot analysis
  umap,                     # Uniform Manifold Approximation Projection
  ggpubr,                   # Simple ggplots with stats
  ggpirate,                 # ggplot version of Pirate plot
  mixOmics,                 # PLS-DA
  conflicted,               # Forces conflicted packages to be used by preferred package
  gridExtra,                # Arrange multiple grobs
  reticulate,               # Run python in R
  tableone,                 # Additional clinical tables support
  lubridate,                # Manipulate date objects
  yingtools2,               # Custom plotting functions
  cowplot,                  # Combine plots together
  epiR,                     # Obtain model performance measures
  pROC,                     # Produce ROC curves
  plotrix,                  # Add table to base R plot
  glmnet,                   # LASSO regression
  forcats,                  # Factor reordering
  ggpmisc,                  # Miscellaneous ggplot helper functions
  doParallel,               # Parallelize functions (for optimizing cutpoints)
  PRISMAstatement,          # Flow table generator
  DiagrammeRsvg,            # Flow table visualizer
  rsvg,                     # Flow table visualizer
  scales,                   # Scale functions for visualizations
  devtools,                 # To load packages from GitHub
  datscience,               # Bibliography manager
  install = F
)

library("gtsummary")  # Clinical tables, pacman had trouble loading this in, so had to go the manual route

conflict_prefer("select", "dplyr")
conflict_prefer("mutate", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("rename", "dplyr")
conflict_prefer("slice", "dplyr")
conflict_prefer("between", "dplyr")
conflict_prefer("annotate", "ggplot2")

devtools::source_url("https://github.com/yingeddi2008/DFIutility/blob/master/getRdpPal.R?raw=TRUE")

# ggplot theme shortcuts
et <- element_text
eb <- element_blank
er <- element_rect
el <- element_line

opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)

`%!in%` <- negate(`%in%`)

# Running python in R
## User must set path to their python library
reticulate::use_python("/usr/bin/python3")

# Customized lefse function
source("./Data/run_lefse2.R")
```

## Import Data
### Load in all Metagenomic, Metabolomic and Clinical Data
```{r, load-in-data, message=FALSE, warning=FALSE}
# 1) Load R image 
# load('./Data/LT_Modeling.RData')

# OR #

# 2) Individually Load R Objects

# Sample lookup
sample_lookup <- readRDS("./Data/sample_lookup.rds")

# First samples from all patients
first_samps <- readRDS("./Data/first_samps_anon.rds")

# Simplified dataframe containing infection information and relative abundance of targeted taxa
peri_matrix_all <- readRDS("./Data/peri_matrix_clin_all.rds")

# Distinct detailed infection data 
peri_criteria_best <- readRDS("./Data/peri_criteria_best_anon.rds")

# All detailed infection data
peri_criteria_all <- readRDS("./Data/peri_criteria_all_anon.rds")

# NCBI taxonomy lookup
tax_lookup <- readRDS("./Data/tax_lookup.rds")

# Complete metaphlan dataframe for all patients and healthy donors
metaphlan_df <- readRDS("./Data/metaphlan_anon.rds")

# Metaphlan dataframe of patient samples
metaphlan_peri_anon <- readRDS("./Data/metaphlan_peri_anon.rds")

# Custom color palette
metaphlan_pal <- getRdpPal(metaphlan_df)

# Qualitative metabolomics
metab_qual_anon <- readRDS("./Data/metab_qual_anon.rds")

# Quantitative metabolomics
metab_quant_anon <- readRDS("./Data/metab_quant_anon.rds")

# Antibiotics data
abx <- readRDS("./Data/abx_anon.rds")

# Demographics data
demo <- readRDS("./Data/demo_anon.rds")
```

## Figures
### Figure 4C: Optimal Cutpoint Analysis: Relative Abundance
```{r, cutpoints, fig.dim = c(16, 12), message=FALSE, warning=FALSE}
# Custom function to avoid any errors during map 
safe_cutpointr <- possibly(.f = cutpointr, otherwise = "Error")

# Calculate the number of cores
no_cores <- detectCores() - 2

# create the cluster for caret to use
cl <- makePSOCKcluster(no_cores)
registerDoParallel(cl)

set.seed(123456)
test_abundance <-
  peri_matrix_all %>%
  mutate(sampleID = as.factor(sampleID)) %>% 
  select(starts_with(c("entero","esch","klebs","citro","rahn","proteus")), -ends_with("abs_abundance")) %>% 
  mutate_if(is.character, as.numeric) %>% 
  mutate(
         enterococcus_infection = ifelse(`Enterococcus faecium` + `Enterococcus faecalis` + `Enterococcus avium` >= 1, 1, 0),
         enterobacterales_infection = ifelse(`Escherichia coli`+ `Klebsiella pneumoniae`+ `Citrobacter freundii`+`Proteus mirabilis` >= 1, 1, 0)) %>%  
  select(-c(`Enterococcus faecium`, `Enterococcus faecalis`, `Enterococcus avium`,
            `Escherichia coli`, `Klebsiella pneumoniae`, `Citrobacter freundii`, `Proteus mirabilis`)) %>% 
  pivot_longer(-c(enterococcus_infection, enterobacterales_infection), names_to = "variable", values_to = "value") %>% 
  pivot_longer(-c(variable, value), names_to = "infection_type", values_to = "infection") %>% 
  mutate(variable = paste0("Input: ", variable, "\nPredict: ", infection_type)) %>% 
  group_by(infection_type, variable) %>% 
  group_map(~ safe_cutpointr(., value, infection, variable, 
                             method = maximize_metric, 
                             metric = youden, 
                             pos_class = 1,
                             boot_runs = 100,
                             allowParallel = TRUE,
                             na.rm = T), .keep = TRUE)

# to get cutpoint object
test_abundance[4][[1]] # ecoc rel abd, ecoc infection
test_abundance[1][[1]] # ebac rel abd, ebacc infection

test_abundance_unnest <-
  test_abundance %>% 
  map_df(as_tibble)


test_abundance_unnest %>% 
  separate(subgroup, into = c("Input", "Predict"), sep = "\n", remove = F) %>% 
  filter(grepl("enterobacterales", Input) & grepl("enterobacterales", Predict)|
         grepl("enterococcus", Input) & grepl("enterococcus", Predict)) %>% 
  group_by(pos_class) %>%
  ungroup() %>% 
  unnest(roc_curve) %>% 
  arrange(desc(AUC)) %>% 
  select(-boot) %>% 
  mutate(
    auc_label = paste0("AUC = ", formatC(round(AUC, 3), digits = 2, format = "f")),
    auc_label = case_when(
      grepl("enterococcus_rel_abundance", Input) ~ paste0(
        auc_label,
        " [",
        formatC(pull(round(boot_ci(x = test_abundance[4][[1]], AUC, alpha = 0.05)[3][1, ], 2)), digits = 2, format = "f"),
        ", ",
        formatC(pull(round(boot_ci(x = test_abundance[4][[1]], AUC, alpha = 0.05)[3][2, ], 2)), digits = 2, format = "f"),
        "]"
      ),
      grepl("enterobacterales_rel_abundance", Input) ~ paste0(
        auc_label,
        " [",
        formatC(pull(round(boot_ci(x = test_abundance[1][[1]], AUC, alpha = 0.05)[3][1, ], 2)), digits = 2, format = "f"),
        ", ",
        formatC(pull(round(boot_ci(x = test_abundance[1][[1]], AUC, alpha = 0.05)[3][2, ], 2)), digits = 2, format = "f"),
        "]"
      )
    ),
    acc_label = paste0("Accuracy = ", formatC(round(acc, 3)*100,  digits = 0, format = "f"), "%"),
    acc_label = case_when(
      grepl("enterococcus_rel_abundance", Input) ~ paste0(
        acc_label,
        " [",
        paste0(formatC(pull(round(boot_ci(x = test_abundance[4][[1]], acc, alpha = 0.05)[3][1, ], 2))*100, digits = 0, format = "f"), "%"),
        ", ",
        paste0(formatC(pull(round(boot_ci(x = test_abundance[4][[1]], acc, alpha = 0.05)[3][2, ], 2))*100, digits = 0, format = "f"), "%"),
        "]"
      ),
      grepl("enterobacterales_rel_abundance", Input) ~ paste0(
        acc_label,
        " [",
        paste0(formatC(pull(round(boot_ci(x = test_abundance[1][[1]], acc, alpha = 0.05)[3][1, ], 2))*100, digits = 0, format = "f"), "%"),
        ", ",
        paste0(formatC(pull(round(boot_ci(x = test_abundance[1][[1]], acc, alpha = 0.05)[3][2, ], 2))*100, digits = 0, format = "f"), "%"),
        "]"
      )
    ),
    sens_label = paste0("Sensitivity = ", formatC(round(sensitivity, 3)*100, digits =0, format = "f"), "%"),
    sens_label = case_when(
      grepl("enterococcus_rel_abundance", Input) ~ paste0(
        sens_label,
        " [",
        paste0(formatC(pull(round(boot_ci(x = test_abundance[4][[1]], sensitivity, alpha = 0.05)[3][1, ], 2))*100, digits = 0, format = "f"), "%"),
        ", ",
        paste0(formatC(pull(round(boot_ci(x = test_abundance[4][[1]], sensitivity, alpha = 0.05)[3][2, ], 2))*100, digits = 0, format = "f"), "%"),
        "]"
      ),
      grepl("enterobacterales_rel_abundance", Input) ~ paste0(
        sens_label,
        " [",
        paste0(formatC(pull(round(boot_ci(x = test_abundance[1][[1]], sensitivity, alpha = 0.05)[3][1, ], 2))*100, digits = 0, format = "f"), "%"),
        ", ",
        paste0(formatC(pull(round(boot_ci(x = test_abundance[1][[1]], sensitivity, alpha = 0.05)[3][2, ], 2))*100, digits = 0, format = "f"), "%"),
        "]"
      )
    ),
    spec_label = paste0("Specificity = ", formatC(round(specificity, 3)*100, digits = 0, format = "f"), "%"),
    spec_label = case_when(
      grepl("enterococcus_rel_abundance", Input) ~ paste0(
        spec_label,
        " [",
        paste0(formatC(pull(round(boot_ci(x = test_abundance[4][[1]], specificity, alpha = 0.05)[3][1, ], 2))*100, digits = 0, format = "f"), "%"),
        ", ",
        paste0(formatC(pull(round(boot_ci(x = test_abundance[4][[1]], specificity, alpha = 0.05)[3][2, ], 2))*100, digits = 0, format = "f"), "%"),
        "]"
      ),
      grepl("enterobacterales_rel_abundance", Input) ~ paste0(
        spec_label,
        " [",
        paste0(formatC(pull(round(boot_ci(x = test_abundance[1][[1]], specificity, alpha = 0.05)[3][1, ], 2))*100, digits = 0, format = "f"), "%"),
        ", ",
        paste0(formatC(pull(round(boot_ci(x = test_abundance[1][[1]], specificity, alpha = 0.05)[3][2, ], 2))*100, digits = 0, format = "f"), "%"),
        "]"
      )
    )
  ) %>% 
  filter(grepl(pattern = "rel_abundance", x = Input)) %>% 
  mutate(subgroup = str_to_title(subgroup),
         subgroup = gsub(pattern = "_", replacement = " ", x = subgroup),
         subgroup = gsub(pattern = "rel abundance", replacement = "Relative Abundance (%)", x = subgroup),
         subgroup2 = ifelse(grepl(x = subgroup, pattern = "enterococcus", ignore.case = TRUE), "Enterococcus", "Enterobacterales"),
         subgroup2 = factor(subgroup2, levels = c("Enterococcus", "Enterobacterales"))) %>% 
  ggplot(aes(x = fpr, y = tpr, color = subgroup2)) +
  geom_line(size = 1.2) +
  geom_text(aes(label = auc_label, x = 0.5, y = 0.13), show.legend = F, hjust = 0) +
  geom_text(aes(label = acc_label, x = 0.5, y = 0.09), show.legend = F, hjust = 0) +
  geom_text(aes(label = spec_label, x = 0.5, y = 0.05), show.legend = F, hjust = 0) +
  geom_text(aes(label = sens_label, x = 0.5, y = 0.01), show.legend = F, hjust = 0) +
  geom_text(aes(label = paste0("Cutpoint = ", round((optimal_cutpoint), digits = 3)*100, "%"), x = 0.5, y = 0.17), hjust = 0, show.legend = F) +
  theme_bw()+
  theme(
    axis.text = et(color = "black", size = 12),
    axis.title = et(color = "black", size = 14),
    legend.text = et(size = 12),
    legend.title = et(size = 14),
    legend.spacing.y = unit(0.5, 'cm'),
    legend.position = "none",
    panel.grid = eb(),
    strip.text = et(size = 14),
    strip.background = eb(),
    panel.border = eb(),
    panel.spacing.y = unit(20, "mm"),
    axis.line.x = el(color = "black")
  ) +
  geom_vline(xintercept = -0.05) +
  geom_hline(yintercept = -0.05) +
  scale_x_continuous(expand = expansion(add = c(0.001,0.05))) +
  scale_y_continuous(expand = expansion(add = c(0.001,0.05))) +
  facet_wrap(~subgroup2, ncol = 1, scales = "fixed") +
  ylab("True Positive Rate\n") +
  xlab("\nFalse Positive Rate") +
  scale_color_manual(values = c("#2dc46b",
                               "red")) +
  guides(color = guide_legend("Groups", 
                              byrow = T,
                              override.aes = list(size = 5)))


ggsave(filename =  "./Results/Figure_4C.pdf", height = 12, width = 7)


# obtain threshold cutpoints for relative abundance
optimal_cutpoint_rel <-
test_abundance_unnest %>% 
  separate(subgroup, into = c("Input", "Predict"), sep = "\n", remove = F) %>% 
  filter(grepl("enterobacterales", Input) & grepl("enterobacterales", Predict)|
         grepl("enterococcus", Input) & grepl("enterococcus", Predict)) %>% 
  group_by(pos_class) %>% 
  filter(grepl("rel", subgroup)) %>% 
  select(subgroup, optimal_cutpoint)
```

### Figure 5A: Enterococcus Expansion: Volcano Analysis
```{r, volcano-analysis-enterococcus-expansion, fig.dim = c(24, 8),warning=FALSE, message=FALSE}
# Heatmap compounds and their categories
heatmap_lookup <- read.csv("./Data/qual_heatmap_lookup.csv", stringsAsFactors = FALSE)

# Build heatmap compound list
heatmap_cmpds <- metab_qual_anon %>% 
  mutate(compound = str_to_title(compound),
         compound = recode(compound,
                          Preq1 = "PreQ1")) %>%
  filter(compound %in% heatmap_lookup$compound) %>% 
  distinct(compound) %>% 
  drop_na()

qual_log2fc_ecoc_expan <- metab_qual_anon %>%
  mutate(compound = ifelse(compound == "isovaleric-acid", "isovalerate", compound),
         compound = str_to_title(compound),
         compound = recode(compound,
                          Preq1 = "PreQ1")) %>% 
  filter(compound %in% heatmap_cmpds$compound) %>%
  left_join(peri_matrix_all %>% select(sampleID, enterococcus_rel_abundance)) %>% 
  drop_na() %>% 
  mutate(enterococcus_expansion = ifelse(enterococcus_rel_abundance >= optimal_cutpoint_rel$optimal_cutpoint[2], 1, 0)) %>% 
  arrange(enterococcus_expansion, sampleID) %>% 
  group_by(compound) %>%
  mutate(enterococcus_expansion_0 = length(mvalue[enterococcus_expansion == "0"]),
         enterococcus_expansion_1 = length(mvalue[enterococcus_expansion == "1"])) %>%
  filter(any(mvalue != 0)) %>% 
  summarise(log2fc_val = log((mean(mvalue[enterococcus_expansion == "0"], na.rm = T)/ mean(mvalue[enterococcus_expansion == "1"], na.rm = T)), base = 2)) # 0 = No Expansion, 1 = Expansion


qual_pval_ecoc_expan <- metab_qual_anon %>%
  mutate(compound = ifelse(compound == "isovaleric-acid", "isovalerate", compound),
         compound = str_to_title(compound),
         compound = recode(compound,
                          Preq1 = "PreQ1")) %>% 
  filter(compound %in% heatmap_cmpds$compound) %>% 
  left_join(peri_matrix_all %>% select(sampleID, enterococcus_rel_abundance)) %>% 
  drop_na() %>% 
  mutate(enterococcus_expansion = ifelse(enterococcus_rel_abundance >= optimal_cutpoint_rel$optimal_cutpoint[2], 1, 0)) %>% 
  arrange(enterococcus_expansion, sampleID) %>% 
  group_by(compound) %>%
  mutate(enterococcus_expansion_0 = length(mvalue[enterococcus_expansion == "0"]),
         enterococcus_expansion_1 = length(mvalue[enterococcus_expansion == "1"])) %>% 
  filter(any(mvalue != 0)) %>% 
  rstatix::wilcox_test(mvalue ~ enterococcus_expansion) %>% 
  rstatix::adjust_pvalue(method = "BH") %>%
  rstatix::add_significance("p.adj")

qual_tot_ecoc_expan <- left_join(qual_log2fc_ecoc_expan, qual_pval_ecoc_expan) %>% 
  column_to_rownames(var = "compound")

# volcano label color
ecoc_expan_volcano_labcol <- qual_tot_ecoc_expan %>% 
  filter(p.adj <= 0.05 & abs(log2fc_val) >= 0.75) %>% 
  mutate(color = ifelse(log2fc_val < 0, "#389458", "gray47"))

# Volcano Plot (adjusted)
set.seed(456)
volcano_adj_ecoc_expan <-
EnhancedVolcano(qual_tot_ecoc_expan,
                lab = rownames(qual_tot_ecoc_expan),
                x = 'log2fc_val',
                y = 'p.adj',
                title = NULL,
                pCutoff = 0.05,
                FCcutoff = 0.75,
                pointSize = 6,
                labSize = 8,
                axisLabSize = 32,
                labCol = ecoc_expan_volcano_labcol$color,
                caption = NULL,
                colAlpha = 0.65,
                col = c("gray75", c("#D4CA15", "#912777", "#1238E3")),
                xlim = c(-2.5, 4),
                ylim = c(0, 8),
                legendPosition = "none",
                legendLabels = c(expression(p.adj > 0.05*";" ~ Log[2] ~ FC < "\u00B1"*0.75),
                          expression(p.adj > 0.05*";" ~ Log[2] ~ FC >= "\u00B1"*0.75),
                          expression(p.adj <= 0.05*";" ~ Log[2] ~ FC < "\u00B1"*0.75),
                          expression(p.adj <= 0.05*";" ~ Log[2] ~ FC >= "\u00B1"*0.75)),
                             legendLabSize = 14,
                boxedLabels = T,
                drawConnectors = T,
                widthConnectors = 0.2,
                arrowheads = F,
                gridlines.minor = F,
                gridlines.major = F,
                max.overlaps = Inf
                ) +
  theme(
    axis.text = et(color = "black"),
    legend.text = et(hjust = 0),
    plot.margin = unit(c(0, 0, 0, 0), "cm")
  ) +
  labs(subtitle = NULL) +
  annotate("segment", x = 0.8, xend = 3, y = 7.95, yend = 7.95, 
           arrow = arrow(), size = 2, color = "gray67") +
  annotate("text", x = 1.65, y = 8.15, label = "No Expansion", size = 9, color = "gray67") +
  annotate("rect", xmin = 0.75, xmax = Inf, ymin = -log(0.05, base = 10), ymax = Inf,
           alpha = .1, fill = "gray87") +
  annotate("segment", x = -0.8, xend = -2.5, y = 7.95, yend = 7.95, 
           arrow = arrow(), size = 2, color = "#389458") +
  annotate("text", x = -1.55, y = 8.15, label = "Expansion", size = 9, color = "#389458") +
  annotate("rect", xmin = -0.75, xmax = -Inf, ymin = -log(0.05, base = 10), ymax = Inf,
           alpha = .1, fill = "#389458") +
  guides(color = guide_legend(nrow = 4),
         shape = guide_legend(nrow = 4)) +
  scale_y_continuous(expand = expansion(add = c(0, 0.15)))

volcano_adj_ecoc_expan

ggsave(plot = volcano_adj_ecoc_expan, filename = "./Results/Figure_5A.pdf", width = 24, height = 11) 
```

### Enterococcus Expansion: Distribution
```{r, qual-npa-enterococcus-expansion-distribution, fig.dim = c(16, 12),warning=FALSE, message=FALSE}
# Using unadjusted p-values for down-selection of metabolites, show distribution of normalized peak area
boxplot_ecoc_expan <- metab_qual_anon %>%
  mutate(compound = ifelse(compound == "isovaleric-acid", "isovalerate", compound),
         compound = str_to_title(compound),
         compound = recode(compound,
                          Preq1 = "PreQ1")) %>% 
  filter(compound %in% heatmap_cmpds$compound) %>%
  left_join(peri_matrix_all %>% select(sampleID, enterococcus_rel_abundance)) %>% 
  drop_na() %>% 
  mutate(enterococcus_expansion = ifelse(enterococcus_rel_abundance >= optimal_cutpoint_rel$optimal_cutpoint[2], 1, 0),
         enterococcus_expansion = ifelse(enterococcus_expansion == 1, "Expansion", "No Expansion")) %>% 
  arrange(enterococcus_expansion, sampleID) %>% 
  group_by(compound) %>%
  mutate(enterococcus_expansion_0 = length(mvalue[enterococcus_expansion == "No Expansion"]),
         enterococcus_expansion_1 = length(mvalue[enterococcus_expansion == "Expansion"])) %>%
  filter(any(mvalue != 0)) %>% 
  right_join(qual_tot_ecoc_expan %>% 
               rownames_to_column(var = "compound") %>% 
               mutate(abs_log2fc_val = abs(log2fc_val)) %>% 
               filter(p <= 0.05, abs_log2fc_val >= 1))

if (nrow(boxplot_ecoc_expan) > 0){
print(ggpar(
  ggboxplot(
    boxplot_ecoc_expan,
    x = "enterococcus_expansion",
    y = "mvalue",
    color = "enterococcus_expansion",
    palette = c("#389458","gray87"),
    ylab = "Normalized Peak Area",
    xlab = "",
    outlier.shape = NA
  ),
  legend.title = "Enterococcus"
) +
  stat_compare_means(label.y.npc = 0.75) +
  facet_wrap(~compound, scales = "free_y") +
  geom_point(data = boxplot_ecoc_expan, aes(x = enterococcus_expansion, y = mvalue, color = enterococcus_expansion), position = position_jitter(width = 0.2), 
             size = 2, alpha = 0.65))

ggsave(filename = "./Results/enterococcus_expansion_boxplot.pdf", width = 24, height = 18)
} else {
  print("no significant observations")
}
```

### taxUMAP Data Preparation
```{r, umap-pre-taxumap, warning=FALSE, message=FALSE, fig.dim = c(16, 12)}
# Custom colors
pirate_colors <- c("#1A49BE", "#3A001E")
pirate_colors2 <- c("#3A001E", "#3A001E", "#3A001E", "#1A49BE")

t_metaphlan <- metaphlan_df %>%
  filter(sampleID %in% first_samps$sampleID| grepl(sampleID, pattern = "hd")) %>% 
  mutate(db = ifelse(grepl(sampleID, pattern = "lt"), "Liver Transplant", "Healthy Donor")) %>% 
  select(sampleID, taxid, db, pctseqs, Total) %>% 
  group_by(sampleID, taxid, pctseqs) %>% 
  slice(1) %>% 
  ungroup() %>% 
  filter(pctseqs >= 0.0001) %>% 
  group_by(sampleID) %>% 
  dplyr::add_count(taxid, name = "totalSp") %>% 
  mutate(sampleID_count = length(unique(sampleID)),
         spPres = totalSp/sampleID_count) %>%
  filter(spPres >= 0.10) %>% 
  select(-c(Total, sampleID_count, spPres, totalSp)) %>% 
  group_by(sampleID) %>% 
  mutate(pctseqs = pctseqs/sum(pctseqs))


t_metaphlan_mat <- t_metaphlan %>%
  distinct() %>% 
  pivot_wider(names_from = "taxid", values_from = "pctseqs", values_fill = 0) %>% 
  column_to_rownames(var = "sampleID") %>% 
  select(-db)


# taxUMAP microbiota table
taxumap_microbiota <- t_metaphlan_mat %>% 
  rownames_to_column(var = "index_column") %>% 
  pivot_longer(-index_column, names_to = "taxid", values_to = "pctseq") %>% 
  mutate(taxid = paste0("taxID", taxid)) %>% 
  pivot_wider(names_from = "taxid", values_from = "pctseq")

write.csv(taxumap_microbiota, "./Results/microbiota_table.csv", row.names = FALSE)

# taxUMAP taxonomy table
taxumap_taxonomy <- t_metaphlan_mat %>% 
  rownames_to_column(var = "index_column") %>% 
  pivot_longer(-index_column, names_to = "taxid", values_to = "pctseq") %>% 
  left_join(tax_lookup %>% mutate(taxid = as.character(taxid))) %>% 
  mutate(taxid = paste0("taxID", taxid)) %>% 
  distinct(Kingdom, Phylum, Class, Order,
           Family, Genus, Species, taxid) %>%
  transmute(OTU = taxid,
            Kingdom,
            Phylum = if_else(Phylum == "", "unclassified", Phylum),
            Class  = if_else(Class == "", "unclassified", Class),
            Order = if_else(Order == "", "unclassified", Order),
            Family = if_else(Family == "", "unclassified", Family),
            Genus = if_else(Genus == "", "unclassified", Genus),
            Genus = gsub(" ","_", Genus),
            Species = if_else(Species == "", "unclassified",Species),
            Species = gsub(" ","_", Species)
            )
write.csv(taxumap_taxonomy, "./Results/taxonomy.csv", row.names = FALSE)
```

### TaxUMAP Analysis
```{python, warning=FALSE, message=FALSE}
import sys

print(sys.path)

# USER MUST SET THEIR PATH TO THEIR LOCALLY INSTALLED TAXUMAP LOCATION
sys.path.insert(0, '/Users/nick/Documents/taxumap/taxumap/')

from taxumap.taxumap_base import Taxumap

# From file
tu = Taxumap(taxonomy='./Results/taxonomy.csv', microbiota_data='./Results/microbiota_table.csv',random_state=456)

# Transform the data (an inplace function)
tu.transform_self(neigh=55)

# Raw embedding dataframe
embedding_df = tu.df_embedding
```

### Figure 1: Shotgun Metagenomics, taxUMAP and Diversity 
```{r, figure-1-relative-abundance-umap-alpha-diversity, fig.dim = c(18, 14), warning=FALSE, message=FALSE}
#### Alpha Diversity ####
# Alpha diversity matrix: Inverse Simpson
alpha_invsim <- vegan::diversity(t_metaphlan_mat, index = "invsimpson") %>%
  as.data.frame() %>% 
  rownames_to_column(var = "sampleID") %>% 
  dplyr::rename("InvSimpson" = ".")

# Alpha Diversity matrix: Shannon
alpha_shannon <- vegan::diversity(t_metaphlan_mat, index = "shannon") %>%
  as.data.frame() %>% 
  rownames_to_column(var = "sampleID") %>% 
  dplyr::rename("Shannon" = ".")

# Alpha Diversity matrix: Observed ASVs
alpha_richness <- vegan::specnumber(t_metaphlan_mat) %>%
  as.data.frame() %>% 
  rownames_to_column(var = "sampleID") %>% 
  dplyr::rename("Richness" = ".")

#### taxUMAP #####
# Find most abundant taxa per sample and plot just that
top_tax <- t_metaphlan %>%
  group_by(sampleID) %>%
  slice_max(pctseqs, n = 1) %>%
  left_join(tax_lookup) %>% 
  mutate(across(everything(), ~ifelse(.=="", NA, as.character(.)))) %>% 
  replace_na(list(Species="unclassified",
                  Genus="unclassified",
                  Family="unclassified",
                  Order="unclassified",
                  Class="unclassified",
                  Phylum="unclassified")) %>%
  mutate(Genus=ifelse(Genus=="unclassified",
                      paste(Family,Genus,sep="\n"),
                      as.character(Genus)),
         pctseqs = as.numeric(pctseqs))


metaphlan_df2 <- t_metaphlan %>% 
  mutate(db = factor(db, levels = c("Liver Transplant", "Healthy Donor"))) %>% 
  left_join(tax_lookup) %>% 
  drop_na(taxid) %>% 
  arrange(Kingdom, Phylum, Class, Order, Family, Genus) %>%
  mutate(Genus = paste0(Phylum,"-",Order,"-", Family, "-",Genus)) %>% 
  left_join(alpha_shannon) %>%
  group_by(sampleID) %>% 
  arrange(Genus) %>% 
  mutate(cum.pct = cumsum(pctseqs), 
         y.text = (cum.pct + c(0, cum.pct[-length(cum.pct)]))/2) %>% 
  ungroup() %>%
  dplyr::select(-cum.pct)

metaphlan_df_sumry <- 
  metaphlan_df2 %>% 
  group_by(sampleID) %>% 
  dplyr::slice(1) %>% 
  ungroup() %>% 
  mutate(healthy_min_shannon = min(Shannon[db == "Healthy Donor"]),
         diversity_group = case_when(
           db == "Healthy Donor" ~ "Healthy Donor",
           Shannon >= healthy_min_shannon ~ "High Diversity",
           TRUE ~ "TBD"
         ),
         lt_med_shannon = median(Shannon[diversity_group == "TBD"], na.rm = TRUE),
         diversity_group = case_when(
           diversity_group %in% c("Healthy Donor", "High Diversity") ~ diversity_group,
           Shannon >= lt_med_shannon ~ "Medium Diversity",
           TRUE ~ "Low Diversity"
         ),
         diversity_group = factor(
           diversity_group,
           levels = c(
             "Low Diversity",
             "Medium Diversity",
             "High Diversity",
             "Healthy Donor"
           )
         ),
         diversity_group_abv = gsub(pattern = " Diversity",
                                    replacement = "",
                                    x = diversity_group),
         diversity_group_abv = factor(diversity_group_abv,
                                      levels = c("Low", "Medium", "High", "Healthy Donor"))) %>% 
  arrange(Genus)


tax_umap_mat_plot <- py$embedding_df %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "sampleID") %>% 
  left_join(t_metaphlan %>% 
              distinct(sampleID, db)) %>%
  left_join(top_tax) %>% 
  ggplot(aes(x = taxumap1, y = taxumap2, color = Genus, shape = db, size = pctseqs*100, alpha = db))+
  geom_point(fill = "black")+
  theme_bw() +
  theme(panel.grid = eb(),
        axis.title = et(color = "black", size = 14),
        axis.text = et(color = "black", size = 12),
        legend.title = et(color = "black", size = 14),
        legend.text = et(color = "black", size = 12),
        legend.position = "top"
        ) + 
  xlab("taxUMAP1") +
  ylab("taxUMAP1") +
  guides(
    shape = guide_legend(
      title = "Cohort",
      override.aes = list(size = 4),
      order = 1,
      nrow = 2,
      title.position = "top",
      title.hjust = 0.5
    ),
    size = guide_legend(
      title = "Relative Abundance (%)",
      order = 2,
      nrow = 2,
      title.position = "top",
      title.hjust = 0.5
    ),
    color = "none",
    fill = "none",
    alpha = "none"
  )+
  scale_color_manual(values = metaphlan_pal)+
  scale_shape_manual(values = c(24, 16))+
  scale_alpha_manual(values = c(1, 0.45))

tax_umap_mat_plot

ggsave(plot = tax_umap_mat_plot, 
            filename = "./Results/Figure_1B.pdf",
            height = 8, width = 8)

#### Alpha Diversity Stats #####
# Obtain stats for alpha diversity
diversity_comps <- list(
  c("Healthy Donor", "High"),
  c("High", "Medium"),
  c("High", "Low"),
  c("Medium", "Low")
  )

alpha_stats <- alpha_invsim %>% 
  left_join(alpha_shannon) %>% 
  left_join(alpha_richness) %>% 
  inner_join(metaphlan_df2 %>%
               left_join(metaphlan_df_sumry %>% select(sampleID, db, diversity_group_abv)) %>% 
               distinct(sampleID, db, diversity_group_abv)
             ) %>% 
  pivot_longer(!c(sampleID, db, diversity_group_abv),names_to = "diversity_metric", values_to = "value") %>% 
  group_by(diversity_metric) %>% 
  rstatix::wilcox_test(value~diversity_group_abv,
                        comparisons = diversity_comps,
                       p.adjust.method = "BH",
                       alternative= "two.sided"
                       ) %>% 
  ungroup()

# Create dataframe for all phylogentic levels of interest
phylo_rel_abd <- t_metaphlan %>% 
  left_join(tax_lookup) %>% 
  inner_join(metaphlan_df2 %>%
               left_join(metaphlan_df_sumry %>% select(sampleID, db, diversity_group_abv)) %>% 
               distinct(sampleID, db, diversity_group_abv)
             ) %>% 
  mutate(Species = paste(Kingdom, Phylum, Class, Order, Family, Genus, Species, sep = "|")) %>% 
  filter(grepl(pattern = "Enterococcus|Enterobacterales|Bacteroidetes|Lachnospiraceae|Oscillospiraceae", x = Species)) %>% 
  mutate(organism = case_when(grepl(pattern = "Enterococcus", x = Species) ~ "Enterococcus",
                              grepl(pattern = "Enterobacterales", x = Species) ~ "Enterobacterales",
                              grepl(pattern = "Bacteroidetes", x = Species) ~ "Bacteroidetes",
                              grepl(pattern = "Lachnospiraceae", x = Species) ~ "Lachnospiraceae",
                              grepl(pattern = "Oscillospiraceae", x = Species) ~ "Oscillospiraceae")) %>% 
  select(sampleID, db, diversity_group_abv, organism, pctseqs) %>% 
  group_by(sampleID, db, diversity_group_abv, organism) %>% 
  summarise(pctseqs = sum(pctseqs)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = organism, values_from = pctseqs, values_fill = 0) %>% 
  pivot_longer(!c(sampleID, db, diversity_group_abv), names_to = "organism", values_to = "pctseqs")

# Obtain stats for all phylogentic levels of interest
rel_abd_alpha_stats <- phylo_rel_abd %>%  
  group_by(organism) %>% 
  rstatix::wilcox_test(pctseqs~diversity_group_abv) %>% 
  bind_rows(alpha_stats) %>% 
  rstatix::adjust_pvalue(method = "BH") %>% 
  mutate(p.adj = ifelse(p.adj < 0.001, 0.001, round(p.adj, 3)))

write.csv(rel_abd_alpha_stats, "./Results/Figure_1_Statistics.csv", row.names = FALSE)

#### Plot Inverse Simpson ####
symnum.args <- list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, Inf), symbols = c("****", "***", "**", "*", "ns"))

diversity_group_colors <- c("#3A001E", "#8A0246", "#C20463", "#1A49BE")

set.seed(456)
gg_alpha_invsim <- alpha_invsim %>% 
  inner_join(t_metaphlan %>% 
               distinct(sampleID, db) %>% 
               select(sampleID, db)
             ) %>% 
  inner_join(metaphlan_df %>%
               left_join(metaphlan_df_sumry %>% select(sampleID, db, diversity_group, diversity_group_abv)) %>% 
               distinct(sampleID, db, diversity_group, diversity_group_abv)
             ) %>% 
  mutate(db = factor(db, levels = c("Liver Transplant", "Healthy Donor"))) %>% 
  ggplot(., aes(x = diversity_group_abv, 
                y = InvSimpson, 
                color = diversity_group_abv, 
                fill = db)) +
  geom_boxplot(outlier.colour = NA,
               alpha = 0.35) +
  geom_jitter(width = 0.2, size = 2.5, alpha = 0.65)+
  stat_compare_means(comparisons = diversity_comps,
                     tip.length = 0.01,
                     step.increase = 0.075,
                     symnum.args = symnum.args,
                     method.args = list(alternative = "two.sided",
                                        exact = FALSE)
                     ) +
  theme_bw() +
  theme(
    panel.grid = eb(),
    axis.title.y = et(size = 14, color = "black"),
    axis.title.x = eb(),
    axis.text = et(size = 12, color = "black"),
    plot.margin = margin(t = 5,  # Top margin
                         r = 5,  # Right margin
                         b = 5,  # Bottom margin
                         l = 5), # Left margin
    panel.border = eb(),
    axis.line = el(color = 'black')
  ) +
  ylab("Alpha Diversity\n(Inverse Simpson)") +
  scale_fill_manual(values = rev(pirate_colors)) +
  scale_color_manual(values = diversity_group_colors) +
  guides(fill = guide_legend("Cohort"),
         color = guide_legend("Diversity Group",
                              override.aes = aes(label = "")))+
  scale_y_continuous(breaks = seq(0,45,5),
                     expand = expansion(mult = c(0.01, 0.1))) +
  coord_cartesian(xlim = c(1.1,3.9))

  
gg_alpha_invsim

pdf(file = "./Results/Figure_1C.pdf", height = 6, width = 7)
gg_alpha_invsim
invisible(dev.off())

#### Plot Shannon Diversity ####
set.seed(456)
gg_alpha_shannon <- alpha_shannon %>% 
  inner_join(t_metaphlan %>% 
               distinct(sampleID, db) %>% 
               select(sampleID, db)
             ) %>% 
  inner_join(metaphlan_df %>%
               left_join(metaphlan_df_sumry %>% select(sampleID, db, diversity_group, diversity_group_abv)) %>% 
               distinct(sampleID, db, diversity_group, diversity_group_abv)
             ) %>% 
  mutate(db = factor(db, levels = c("Liver Transplant", "Healthy Donor"))) %>% 
  ggplot(., aes(x = diversity_group_abv, 
                y = Shannon, 
                color = diversity_group_abv, 
                fill = db)) +
  geom_boxplot(outlier.colour = NA,
               alpha = 0.35) +
  geom_jitter(width = 0.2, size = 2.5, alpha = 0.65)+
  stat_compare_means(comparisons = diversity_comps,
                     tip.length = 0.01,
                     step.increase = 0.075,
                     symnum.args = symnum.args,
                     method.args = list(alternative = "two.sided",
                                        exact = FALSE)
                     ) + 
  theme_bw() +
  theme(
    panel.grid = eb(),
    axis.title.y = et(size = 14, color = "black"),
    axis.title.x = eb(),
    axis.text = et(size = 12, color = "black"),
    plot.margin = margin(t = 5,  # Top margin
                         r = 5,  # Right margin
                         b = 5,  # Bottom margin
                         l = 5), # Left margin
    panel.border = eb(),
    axis.line = el(color = 'black')
  ) +
  ylab("Alpha Diversity\n(Shannon)") +
  scale_fill_manual(values = rev(pirate_colors)) +
  scale_color_manual(values = diversity_group_colors) +
  guides(fill = guide_legend("Cohort"),
         color = guide_legend("Diversity Group",
                              override.aes = aes(label = "")))+
  scale_y_continuous(breaks = seq(0,6.5,1),
                     expand = expansion(mult = c(0.01, 0.1))) +
  coord_cartesian(xlim = c(1.1,3.9))
  
  
gg_alpha_shannon

pdf(file = "./Results/Figure_1D.pdf", height = 6, width = 7)
gg_alpha_shannon
invisible(dev.off())

#### Plot Species Richness ####
set.seed(456)
gg_alpha_richness <- alpha_richness %>% 
  inner_join(t_metaphlan %>% 
               distinct(sampleID, db) %>% 
               select(sampleID, db)
             ) %>% 
  inner_join(metaphlan_df %>%
               left_join(metaphlan_df_sumry %>% select(sampleID, db, diversity_group, diversity_group_abv)) %>% 
               distinct(sampleID, db, diversity_group, diversity_group_abv)
             ) %>% 
  mutate(db = factor(db, levels = c("Liver Transplant", "Healthy Donor"))) %>% 
  ggplot(., aes(x = diversity_group_abv, 
                y = Richness, 
                color = diversity_group_abv, 
                fill = db)) +
  geom_boxplot(outlier.colour = NA,
               alpha = 0.35) +
  geom_jitter(width = 0.2, size = 2.5, alpha = 0.65)+
  stat_compare_means(comparisons = diversity_comps,
                     tip.length = 0.01,
                     label.y = c(145, 160, 150, 110),
                     symnum.args = symnum.args,
                     method.args = list(alternative = "two.sided",
                                        exact = FALSE)
                     ) +
  theme_bw() +
  theme(
    panel.grid = eb(),
    axis.title.y = et(size = 14, color = "black"),
    axis.title.x = eb(),
    axis.text = et(size = 12, color = "black"),
    plot.margin = margin(t = 5,  # Top margin
                         r = 5,  # Right margin
                         b = 5,  # Bottom margin
                         l = 5), # Left margin
    panel.border = eb(),
    axis.line = el(color = 'black')
  ) +
  ylab("Alpha Diversity\n(Richness)") +
  scale_fill_manual(values = rev(pirate_colors)) +
  scale_color_manual(values = diversity_group_colors) +
  guides(fill = guide_legend("Cohort"),
         color = guide_legend("Diversity Group",
                              override.aes = aes(label = "")))+
  scale_y_continuous(breaks = seq(0,180,50),
                     expand = expansion(mult = c(0.01, 0.035))) +
  coord_cartesian(xlim = c(1.1,3.9))
  
  
gg_alpha_richness

pdf(file = "./Results/Figure_1E.pdf", height = 6, width = 7)
gg_alpha_richness
invisible(dev.off())

#### Plot Lachnospiraceae ####
set.seed(456)
gg_lach_rel_abd <- phylo_rel_abd %>% 
  filter(organism == "Lachnospiraceae") %>% 
  group_by(organism) %>% 
  mutate(db = factor(db, levels = c("Liver Transplant", "Healthy Donor"))) %>% 
  ggplot(., aes(x = diversity_group_abv, 
                y = pctseqs, 
                color = diversity_group_abv, 
                fill = db)) +
  geom_boxplot(outlier.colour = NA,
               alpha = 0.35) +
  geom_jitter(width = 0.2, size = 2.5, alpha = 0.65)+
  stat_compare_means(comparisons = diversity_comps,
                     tip.length = 0.01,
                     symnum.args = symnum.args,
                     method.args = list(alternative = "two.sided",
                                        exact = FALSE),
                     label.y = c(0.65, 0.71, 0.76, 0.82)
                     ) +
  theme_bw() +
  theme(
    panel.grid = eb(),
    axis.title.y = et(size = 14, color = "black"),
    axis.title.x = eb(),
    axis.text = et(size = 12, color = "black"),
    plot.margin = margin(t = 5,  # Top margin
                         r = 5,  # Right margin
                         b = 5,  # Bottom margin
                         l = 5), # Left margin
    panel.border = eb(),
    axis.line = el(color = 'black')
  ) +
  ylab(~atop(paste(italic("Lachnospiraceae")), paste("MetaPhlAn4 Relative Abundance"))) +
  scale_fill_manual(values = rev(pirate_colors)) +
  scale_color_manual(values = diversity_group_colors) +
  guides(fill = guide_legend("Cohort"),
         color = guide_legend("Diversity Group",
                              override.aes = aes(label = "")))+
  scale_y_continuous(breaks = seq(0,1,0.1),
                     expand = expansion(mult = c(0.01, 0.035)),
                     labels = scales::percent_format(accuracy = 1)) +
  coord_cartesian(xlim = c(1.1,3.9))
  
  
gg_lach_rel_abd

pdf(file = "./Results/Figure_1F.pdf", height = 6, width = 7)
gg_lach_rel_abd
invisible(dev.off())

#### Plot Enterococcus ####
set.seed(456)
gg_ecoc_rel_abd <- phylo_rel_abd %>% 
  filter(organism == "Enterococcus") %>% 
  group_by(organism) %>% 
  mutate(db = factor(db, levels = c("Liver Transplant", "Healthy Donor"))) %>% 
  ggplot(., aes(x = diversity_group_abv, 
                y = pctseqs, 
                color = diversity_group_abv, 
                fill = db)) +
  geom_boxplot(outlier.colour = NA,
               alpha = 0.35) +
  geom_jitter(width = 0.2, size = 2.5, alpha = 0.65)+
  stat_compare_means(comparisons = diversity_comps,
                     tip.length = 0.01,
                     symnum.args = symnum.args,
                     method.args = list(alternative = "two.sided",
                                        exact = FALSE),
                    
                     label.y = c(0.20, 0.75, 0.985, 1.05)
                     ) +
  theme_bw() +
  theme(
    panel.grid = eb(),
    axis.title.y = et(size = 14, color = "black"),
    axis.title.x = eb(),
    axis.text = et(size = 12, color = "black"),
    plot.margin = margin(t = 5,  # Top margin
                         r = 5,  # Right margin
                         b = 5,  # Bottom margin
                         l = 5), # Left margin
    panel.border = eb(),
    axis.line = el(color = 'black')
  ) +
  ylab(~atop(paste(italic("Enterococcus")), paste("MetaPhlAn4 Relative Abundance"))) +
  scale_fill_manual(values = rev(pirate_colors)) +
  scale_color_manual(values = diversity_group_colors) +
  guides(fill = guide_legend("Cohort"),
         color = guide_legend("Diversity Group",
                              override.aes = aes(label = "")))+
  scale_y_continuous(breaks = seq(0,1,0.1),
                     expand = expansion(mult = c(0.01, 0.035)),
                     labels = scales::percent_format(accuracy = 1)) +
  coord_cartesian(xlim = c(1.1,3.9))
  
  
gg_ecoc_rel_abd

pdf(file = "./Results/Figure_1G.pdf", height = 6, width = 7)
gg_ecoc_rel_abd
invisible(dev.off())

#### Plot Enterobacterales #####
set.seed(456)
gg_ebac_rel_abd <- phylo_rel_abd %>% 
  filter(organism == "Enterobacterales") %>% 
  group_by(organism) %>% 
  mutate(db = factor(db, levels = c("Liver Transplant", "Healthy Donor"))) %>% 
  ggplot(., aes(x = diversity_group_abv, 
                y = pctseqs, 
                color = diversity_group_abv, 
                fill = db)) +
  geom_boxplot(outlier.colour = NA,
               alpha = 0.35) +
  geom_jitter(width = 0.2, size = 2.5, alpha = 0.65)+
  stat_compare_means(comparisons = diversity_comps,
                     tip.length = 0.01,
                     symnum.args = symnum.args,
                     method.args = list(alternative = "two.sided",
                                        exact = FALSE),
                     label.y = c(0.20, 0.75, 0.985, 1.05)
                     ) +
  theme_bw() +
  theme(
    panel.grid = eb(),
    axis.title.y = et(size = 14, color = "black"),
    axis.title.x = eb(),
    axis.text = et(size = 12, color = "black"),
    plot.margin = margin(t = 5,  # Top margin
                         r = 5,  # Right margin
                         b = 5,  # Bottom margin
                         l = 5), # Left margin
    panel.border = eb(),
    axis.line = el(color = 'black')
  ) +
  ylab(~atop(paste(italic("Enterobacterales")), paste("MetaPhlAn4 Relative Abundance"))) +
  scale_fill_manual(values = rev(pirate_colors)) +
  scale_color_manual(values = diversity_group_colors) +
  guides(fill = guide_legend("Cohort"),
         color = guide_legend("Diversity Group",
                              override.aes = aes(label = "")))+
  scale_y_continuous(breaks = seq(0,1,0.1),
                     expand = expansion(mult = c(0.01, 0.035)),
                     labels = scales::percent_format(accuracy = 1)) +
  coord_cartesian(xlim = c(1.1,3.9))
  
  
gg_ebac_rel_abd

pdf(file = "./Results/Figure_1H.pdf", height = 6, width = 7)
gg_ebac_rel_abd
invisible(dev.off())

#### Plot Bacteroidetes ####
set.seed(456)
gg_bact_rel_abd <- phylo_rel_abd %>% 
  filter(organism == "Bacteroidetes") %>% 
  group_by(organism) %>% 
  mutate(db = factor(db, levels = c("Liver Transplant", "Healthy Donor"))) %>% 
  ggplot(., aes(x = diversity_group_abv, 
                y = pctseqs, 
                color = diversity_group_abv, 
                fill = db)) +
  geom_boxplot(outlier.colour = NA,
               alpha = 0.35) +
  geom_jitter(width = 0.2, size = 2.5, alpha = 0.65)+
  stat_compare_means(comparisons = diversity_comps,
                     tip.length = 0.01,
                     symnum.args = symnum.args,
                     method.args = list(alternative = "two.sided",
                                        exact = FALSE),
                     label.y = c(0.65, 0.98, 0.92, 1.02)
                     ) + 
  theme_bw() +
  theme(
    panel.grid = eb(),
    axis.title.y = et(size = 14, color = "black"),
    axis.title.x = eb(),
    axis.text = et(size = 12, color = "black"),
    plot.margin = margin(t = 5,  # Top margin
                         r = 5,  # Right margin
                         b = 5,  # Bottom margin
                         l = 5), # Left margin
    panel.border = eb(),
    axis.line = el(color = 'black')
  ) +
  ylab(~atop(paste(italic("Bacteroidetes")), paste("MetaPhlAn4 Relative Abundance"))) +
  scale_fill_manual(values = rev(pirate_colors)) +
  scale_color_manual(values = diversity_group_colors) +
  guides(fill = guide_legend("Cohort"),
         color = guide_legend("Diversity Group",
                              override.aes = aes(label = "")))+
  scale_y_continuous(breaks = seq(0,1,0.1),
                     expand = expansion(mult = c(0.01, 0.035)),
                     labels = scales::percent_format(accuracy = 1)) +
  coord_cartesian(xlim = c(1.1,3.9))
  
  
gg_bact_rel_abd

pdf(file = "./Results/Figure_1I.pdf", height = 6, width = 7)
gg_bact_rel_abd
invisible(dev.off())

#### Plot Oscillospiraceae ####
set.seed(456)
gg_oscl_rel_abd <- phylo_rel_abd %>% 
  filter(organism == "Oscillospiraceae") %>% 
  group_by(organism) %>% 
  mutate(db = factor(db, levels = c("Liver Transplant", "Healthy Donor"))) %>% 
  ggplot(., aes(x = diversity_group_abv, 
                y = pctseqs, 
                color = diversity_group_abv, 
                fill = db)) +
  geom_boxplot(outlier.colour = NA,
               alpha = 0.35) +
  geom_jitter(width = 0.2, size = 2.5, alpha = 0.65)+
  stat_compare_means(comparisons = diversity_comps,
                     tip.length = 0.01,
                     symnum.args = symnum.args,
                     method.args = list(alternative = "two.sided",
                                        exact = FALSE),
                     label.y = c(0.33, 0.52, 0.59, 0.66)
                     ) +
  theme_bw() +
  theme(
    panel.grid = eb(),
    axis.title.y = et(size = 14, color = "black"),
    axis.title.x = eb(),
    axis.text = et(size = 12, color = "black"),
    plot.margin = margin(t = 5,  # Top margin
                         r = 5,  # Right margin
                         b = 5,  # Bottom margin
                         l = 5), # Left margin
    panel.border = eb(),
    axis.line = el(color = 'black')
  ) +
  ylab(~atop(paste(italic("Oscillospiraceae")), paste("MetaPhlAn4 Relative Abundance"))) +
   scale_fill_manual(values = rev(pirate_colors)) +
  scale_color_manual(values = diversity_group_colors) +
  guides(fill = guide_legend("Cohort"),
         color = guide_legend("Diversity Group",
                              override.aes = aes(label = "")))+
  scale_y_continuous(breaks = seq(0,1,0.1),
                     expand = expansion(mult = c(0.01, 0.035)),
                     labels = scales::percent_format(accuracy = 1)) +
  coord_cartesian(xlim = c(1.1,3.9))
  
  
gg_oscl_rel_abd

pdf(file = "./Results/Figure_1J.pdf", height = 6, width = 7)
gg_oscl_rel_abd
invisible(dev.off())

#### Plot Metaphlan Relative Abundance ####
# Figure 1A-MetaPhlAn4 Taxonomy
# Load legend
tax_legend <- magick::image_read_pdf("./Data/legend.v2.pdf")

gg_tax_legend <- cowplot::ggdraw() + cowplot::draw_image(tax_legend)

metaphlan_pal2 <- getRdpPal(metaphlan_df2)

gg_metaphlan <- metaphlan_df2 %>%
  left_join(metaphlan_df_sumry %>% select(sampleID, diversity_group)) %>% 
  ungroup() %>% 
  mutate(Genus = factor(Genus, levels = unique(Genus))) %>% 
  group_by(sampleID) %>% 
  arrange(Genus) %>% 
  ggplot(aes(x=reorder(sampleID, Shannon),y=pctseqs)) +
  geom_bar(stat="identity",aes(fill=Genus), width = 0.9) +
  scale_fill_manual(values = metaphlan_pal2) +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x=eb(),
        axis.ticks.x=eb(),
        strip.text.x= et(angle=0,size=14),
        strip.background = eb(),
        axis.title.y = et(color = "black", size = 14),
        axis.text.y = et(color = "black", size = 12),
        panel.spacing = unit(0.5, "lines"),
        plot.margin = margin(t = 5,
                             r = 5, 
                             b = 0, 
                             l = 5)) +
  facet_grid(. ~diversity_group, scales = "free", space = "free")+
  scale_y_continuous(expand = expansion(mult = c(0.005,0.005)),
                     labels = scales::percent_format(accuracy = 1)) +
  ylab("MetaPhlAn4 Relative Abundance") +
  xlab("")


# Color facets
gg_metaphlan_grob <- ggplot_gtable(ggplot_build(gg_metaphlan))
strip_both <- which(grepl('strip-', gg_metaphlan_grob$layout$name))
fills <- diversity_group_colors
k <- 1
for (i in strip_both) {
  l <- which(grepl('titleGrob', gg_metaphlan_grob$grobs[[i]]$grobs[[1]]$childrenOrder))
  gg_metaphlan_grob$grobs[[i]]$grobs[[1]]$children[[l]]$children[[1]]$gp$col <- fills[k]
  k <- k+1
}

gg_shannon <- metaphlan_df_sumry %>% 
  ggplot(aes(x=reorder(sampleID, Shannon), y = Shannon)) +
  geom_bar(stat="identity",aes(fill=diversity_group_abv), width = 0.9) +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = eb(),
        axis.title.x = eb(),
        axis.ticks.x = eb(),
        strip.text = eb(),
        strip.background = er(fill = "white"),
        axis.title.y = et(color = "black", size = 14),
        axis.text.y = et(color = "black", size = 12),
        panel.spacing = unit(0.5, "lines"),
        plot.margin = margin(t = 0,
                             r = 5, 
                             b = 0, 
                             l = 5),
        panel.grid = eb()) +
  scale_fill_manual(values = diversity_group_colors) +
  facet_grid(. ~diversity_group, scales = "free", space = "free")

gg_metaphlan_shannon <-
plot_grid(
  gg_metaphlan_grob,
  gg_shannon,
  axis = "lb",
  align = "hv",
  nrow = 2,
  rel_heights = c(1, 0.15)
)

pdf(file = "./Results/Figure_1A.pdf", width = 12.25, height = 8)
gg_metaphlan_shannon 
invisible(dev.off())

#### Combine all into a single figure 1 Start ####
alpha_org_plot <- plot_grid(
  gg_alpha_invsim + theme(legend.position = "none", 
                          axis.text.x = eb(), 
                          axis.title.x = eb(),
                          plot.margin = unit(c(0.1, 0, 0.15, 0), "cm")),
  gg_alpha_shannon + theme(legend.position = "none", 
                           axis.text.x = eb(), 
                           axis.title.x = eb(),
                           plot.margin = unit(c(0.1, 0, 0.15, 0), "cm")),
  gg_alpha_richness+ theme(legend.position = "none", 
                           axis.text.x = eb(),
                           axis.title.x = eb(),
                           plot.margin = unit(c(0.1, 0, 0.15, 0), "cm")),
  gg_lach_rel_abd + theme(axis.text.x = eb(),
                          axis.title.x = eb(),
                          plot.margin = unit(c(0.1, 0, 0.15, 0), "cm")),
  gg_ecoc_rel_abd + theme(legend.position = "none", 
                          axis.text.x = et(angle = 45, hjust = 0.85, vjust = 0.85), 
                          plot.margin = unit(c(0.05, 0, 0, 0), "cm")),
  gg_ebac_rel_abd + theme(legend.position = "none", 
                          axis.text.x = et(angle = 45, hjust = 0.85, vjust = 0.85), 
                          plot.margin = unit(c(0.05, 0, 0, 0), "cm")),
  gg_bact_rel_abd + theme(legend.position = "none", 
                          axis.text.x = et(angle = 45, hjust = 0.85, vjust = 0.85), 
                          plot.margin = unit(c(0.05, 0, 0, 0), "cm")),
  gg_oscl_rel_abd + theme(axis.text.x = et(angle = 45, hjust = 0.85, vjust = 0.85), 
                          plot.margin = unit(c(0.05, 0, 0, 0), "cm")),
  nrow = 2,
  axis = "lb",
  align = "v",
  rel_widths = c(1, 1, 1, 1.3),
  rel_heights = c(1,1.2)
)

# This is useful to figure out the general layout of the plots
gs <- lapply(1:4, function(ii)
  grobTree(rectGrob(gp=gpar(fill=ii, alpha=0.5)), textGrob(ii)))

grid.arrange(grobs=gs, ncol=4,
               top="top label", bottom="bottom\nlabel",
               left="left label", right="right label")

lay <- rbind(c(1,1,1,1,1,3,3,2),
             c(1,1,1,1,1,3,3,2),
             c(4,4,4,4,4,4,NA,NA),
             c(4,4,4,4,4,4,4,4),
             c(4,4,4,4,4,4,4,4))

grid.arrange(grobs = gs, layout_matrix = lay)

pdf(file = "./Results/Figure_1.pdf", height = 16, width = 20)

grid.arrange(
  gg_metaphlan_shannon,                        # 1
  gg_tax_legend,                               # 3
  tax_umap_mat_plot + 
    theme(
      legend.text = et(size = 10),
      legend.title = et(size = 12),
      plot.margin = margin(
        t = 5,  # Top margin
        r = 0,  # Right margin
        b = 75, # Bottom margin
        l = 5   # Left margin
      )
    ),                                      # 2
  alpha_org_plot,                           # 4
  layout_matrix = lay
)

invisible(dev.off())


grid.arrange(
  gg_metaphlan_shannon,                        # 1
  gg_tax_legend,                               # 2
  tax_umap_mat_plot + 
    theme(
      legend.text = et(size = 10),
      legend.title = et(size = 12),
      plot.margin = margin(
         t = 5,  # Top margin
        r = 0,  # Right margin
        b = 75, # Bottom margin
        l = 5   # Left margin
      )
    ),                                      # 3
  alpha_org_plot,                           # 4
  layout_matrix = lay
)

#### Combine all into a single figure 1 End ####
```

### Figure 2: Qualitative Metabolomics Heatmap
```{r, figure-2-qualitative-heatmap, fig.dim = c(26,15), message=FALSE, warning=FALSE}
heatmap_data <- 
  t_metaphlan %>% 
  drop_na(taxid) %>% 
  select(sampleID) %>% 
  group_by(sampleID) %>% 
  dplyr::slice(1) %>% 
  mutate(db = ifelse(grepl(sampleID, pattern = "lt"), "Liver Transplant", "Healthy Donor")) %>%
  left_join(metab_qual_anon) %>% 
  mutate(compound = ifelse(compound == "isovaleric-acid", "isovalerate", compound),
         compound = str_to_title(compound),
         compound = recode(compound,
                          Preq1 = "PreQ1")) %>% 
  filter(compound %in% heatmap_cmpds$compound|is.na(compound)) %>% 
  group_by(compound) %>% 
  mutate(median_val = ifelse(median(mvalue, na.rm = TRUE) == 0, min(mvalue[mvalue > 0], na.rm = TRUE)/10, median(mvalue, na.rm = TRUE)),
         heatmap_val = ifelse(log(mvalue/ median_val, base = 2) == -Inf, 0, log(mvalue/ median_val, base = 2))) %>% 
  ungroup() %>% 
  select(-c(mvalue, median_val)) %>% 
  group_by(sampleID, compound) %>% 
  slice_max(heatmap_val, with_ties = F, n = 1) %>% 
  ungroup() %>%
  select(-db) %>% 
  left_join(metaphlan_df2 %>%
               left_join(metaphlan_df_sumry %>% select(sampleID, db, diversity_group_abv)) %>%
               distinct(sampleID, db, diversity_group_abv), by = "sampleID") %>% 
  group_by(sampleID, compound) %>% 
  slice_max(heatmap_val, with_ties = F, n = 1) %>% 
  left_join(alpha_shannon) %>% 
  group_by(db) %>% 
  arrange(db, Shannon) %>% 
  ungroup()

# Stats
heatmap_pvals <-
  heatmap_data %>%
  filter(diversity_group_abv != "Healthy Donor") %>% 
  drop_na(compound) %>%
  group_by(compound) %>%
  rstatix::kruskal_test(heatmap_val~diversity_group_abv) %>%
  rstatix::adjust_pvalue(method = "BH") %>%
  select(compound, statistic, p, p.adj) 

heatmap_labels <- 
  heatmap_data %>% 
  mutate(compound = str_to_title(compound)) %>% 
  left_join(heatmap_lookup) %>%
  left_join(heatmap_pvals %>% group_by(compound, p.adj, p) %>% slice(1)) %>%
  arrange(class, subclass, p) %>%
  arrange(class, subclass) %>% 
  pivot_wider(c(diversity_group_abv, sampleID, db, Shannon), names_from = "compound", values_from = "heatmap_val", values_fn = mean) %>% 
  group_by(db) %>% 
  arrange(db, Shannon) %>% 
  ungroup() %>% 
  distinct(sampleID, .keep_all = TRUE) %>% 
  select(diversity_group_abv) %>% 
  mutate(
    diversity_group_abv = as.character(diversity_group_abv),
    diversity_group_abv = ifelse(
      grepl(pattern = "Healthy", as.character(diversity_group_abv)),
      diversity_group_abv,
      paste(diversity_group_abv, "Diversity")
    ),
    diversity_group_abv = factor(diversity_group_abv, levels = c("Low Diversity", "Medium Diversity",
                                                                 "High Diversity", "Healthy Donor"))
  )
  
# Build heatmap compound order (row order) and compound class (row slice)
heatmap_order <- 
  heatmap_data %>% 
  ungroup() %>% 
  filter(compound %in% heatmap_cmpds$compound) %>% 
  distinct(compound) %>% 
  drop_na() %>% 
  left_join(heatmap_lookup) %>% 
  left_join(heatmap_pvals %>% group_by(compound, p.adj, p) %>% slice(1)) %>%
  ungroup() %>% 
  mutate(class = factor(
    class,
    levels = c(
      "Fatty Acid",                   # 1
      "Dicarboxylic Acid",            # 2
      "Amino Acid",                   # 3
      "Bile Acid",                    # 4
      "Indole",                       # 5
      "Phenolic Aromatic",            # 6
      "Kynurine Pathway",             # 7
      "Vitamin"                       # 8
    )
  ),
  subclass = factor(
    subclass,
    levels = c(
      "Short-Chain Fatty Acid",       # 1
      "Branched-Chain Fatty Acid",    # 2
      "Aminated Fatty Acid",          # 3
      "Long-Chain Fatty Acid",        # 4
      "Dicarboxylic Acid",            # 5
      "Amino Acid",                   # 6
      "Primary Bile Acid",            # 7
      "Secondary Bile Acid",          # 8
      "Conjugated Bile Acid",         # 9
      "Indole",                       # 10
      "Phenolic Aromatic",            # 11
      "Kynurine Pathway",             # 12
      "Vitamin"                       # 13
    )
  )) %>% 
  arrange(class, subclass, p, compound) %>%
  select(class, subclass, p, compound)
  
# Build heatmap patient order following same order as gg_metaphlan plot
heatmap_column_order <- heatmap_data %>% 
  group_by(patientID) %>% 
  dplyr::slice(1) %>% 
  left_join(alpha_shannon) %>% 
  group_by(db) %>% 
  arrange(db, Shannon) %>%  
  select(patientID) %>% 
  distinct(patientID) %>% 
  pull(patientID)

# P-value legend color
pvalue_col_fun = colorRamp2(c(0, 0.045), c("#75C236", "#E3E4E6"))

# Create heatmap for adjusted p-values
pvalue_adj <-
heatmap_pvals %>% 
  group_by(compound, p.adj, p) %>% 
  dplyr::slice(1) %>% 
  ungroup() %>% 
  left_join(heatmap_lookup) %>%
  mutate(class = factor(
    class,
    levels = c(
      "Fatty Acid",                   # 1
      "Dicarboxylic Acid",            # 2
      "Amino Acid",                   # 3
      "Bile Acid",                    # 4
      "Indole",                       # 5
      "Phenolic Aromatic",            # 6
      "Kynurine Pathway",             # 7
      "Vitamin"                       # 8
    )
  ),
  subclass = factor(
    subclass,
    levels = c(
      "Short-Chain Fatty Acid",       # 1
      "Branched-Chain Fatty Acid",    # 2
      "Aminated Fatty Acid",          # 3
      "Long-Chain Fatty Acid",        # 4
      "Dicarboxylic Acid",            # 5
      "Amino Acid",                   # 6
      "Primary Bile Acid",            # 7
      "Secondary Bile Acid",          # 8
      "Conjugated Bile Acid",         # 9
      "Indole",                       # 10
      "Phenolic Aromatic",            # 11
      "Kynurine Pathway",             # 12
      "Vitamin"                       # 13
    )
  )) %>% 
  arrange(class, subclass, p.adj, compound) %>%
  select(class, subclass, p.adj, compound) %>% 
  column_to_rownames(var = "compound") %>% 
  select(`Adjusted p-value` = p.adj) %>% 
  as.matrix() %>% 
Heatmap(name = "Significance (adjusted p-value)",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        col = pvalue_col_fun,        
        column_title = "KW Test",
        column_title_side = "bottom",
        column_title_rot = 90,
        show_column_names = FALSE,
        column_names_side = "top",
        row_names_gp = gpar(fontsize = 8),
        rect_gp = gpar(col = "black", lwd = 0.2),
        row_gap = unit(3.5, "mm"),
        row_names_side = c("right"),
        row_order = heatmap_order$compound,
        row_split = heatmap_order$subclass,
        show_row_names = FALSE,
        row_title_rot = 0,
        heatmap_legend_param = list(at = seq(0.05, 0, -0.01)),
        width = unit(0.1, "in")
)

# Heatmap legend color
col_fun <- colorRamp2(breaks = c(-5, 0, 5), colors = c("#00aaad", "white", "#ad003a"))

# Global parameter for annotation 
# ht_opt$COLUMN_ANNO_PADDING <- unit(2.5, "mm")

gg_metab_heatmap <-
  heatmap_data %>% 
  mutate(compound = str_to_title(compound),
         compound = recode(compound,
                          Preq1 = "PreQ1")) %>%
  left_join(heatmap_lookup) %>% 
  left_join(heatmap_pvals %>% group_by(compound, p.adj, p) %>% slice(1) %>% select(compound, p.adj, p), by = "compound") %>%
  mutate(class = factor(
    class,
    levels = c(
      "Fatty Acid",                   # 1
      "Dicarboxylic Acid",            # 2
      "Amino Acid",                   # 3
      "Bile Acid",                    # 4
      "Indole",                       # 5
      "Phenolic Aromatic",            # 6
      "Kynurine Pathway",             # 7
      "Vitamin"                       # 8
    )
  ),
  subclass = factor(
    subclass,
    levels = c(
      "Short-Chain Fatty Acid",       # 1
      "Branched-Chain Fatty Acid",    # 2
      "Aminated Fatty Acid",          # 3
      "Long-Chain Fatty Acid",        # 4
      "Dicarboxylic Acid",            # 5
      "Amino Acid",                   # 6
      "Primary Bile Acid",            # 7
      "Secondary Bile Acid",          # 8
      "Conjugated Bile Acid",         # 9
      "Indole",                       # 10
      "Phenolic Aromatic",            # 11
      "Kynurine Pathway",             # 12
      "Vitamin"                       # 13
    )
  )) %>%
  mutate(
    diversity_group_abv = as.character(diversity_group_abv),
    diversity_group_abv = ifelse(
      grepl(pattern = "Healthy", as.character(diversity_group_abv)),
      diversity_group_abv,
      paste(diversity_group_abv, "Diversity")
    ),
    diversity_group_abv = factor(diversity_group_abv, levels = c("Low Diversity", "Medium Diversity",
                                                                 "High Diversity", "Healthy Donor"))
  ) %>% 
  arrange(class, subclass, p.adj, compound) %>%
  pivot_wider(c(diversity_group_abv, db, patientID, Shannon), names_from = compound, values_from = heatmap_val) %>%
  group_by(patientID) %>% 
  arrange(patientID) %>% 
  ungroup() %>% 
  select(-`NA`) %>%
  distinct(patientID, .keep_all = TRUE) %>%
  group_by(diversity_group_abv) %>% 
  arrange(db, Shannon) %>% 
  ungroup() %>% 
  select(-Shannon, -db, -diversity_group_abv) %>% 
  column_to_rownames("patientID") %>%
  as.matrix() %>% 
  t() %>%
  Heatmap(
    name = "Fold Change (log2)",
    col = col_fun,
    na_col = "grey83",
    rect_gp = gpar(col = "grey40", lwd = 1.5),
    column_names_gp = grid::gpar(fontsize = 8),
    column_gap = unit(2.5, "mm"),
    column_split = heatmap_labels,
    column_order = heatmap_column_order,
    column_title_gp = gpar(fontsize = 14),
    column_title_rot = 0,
    cluster_columns = FALSE,
    show_column_names = TRUE,
    show_column_dend = FALSE,
    row_names_gp = gpar(fontsize = 8),
    row_gap = unit(3.5, "mm"),
    row_names_side = c("left"),
    row_order = heatmap_order$compound,
    row_split = heatmap_order$subclass,
    row_title_rot = 0,
    cluster_rows = FALSE,
    show_row_dend = FALSE,
    heatmap_height = unit(13, "in"),
    heatmap_width =  unit(19, "in")
  )

gg_metab_heatmap_tot <- gg_metab_heatmap + pvalue_adj
gg_metab_heatmap_tot


pdf(file = "./Results/Figure_2.pdf", height = 15, width = 26, onefile = F)
gg_metab_heatmap_tot
invisible(dev.off())
```

### Figure 3: Quant Metabolomics-Healthy Donor vs LT Patients
```{r, figure-3-quant-metabolomics, warning=FALSE, message=FALSE, fig.dim = c(16, 12)}
# Conversions of compounds
metab_conversions <- data.frame(
  compound = c(
    "Taurocholic Acid",
    "Glycocholic Acid",
    "Cholic Acid",
    "3-Oxolithocholic Acid",
    "Alloisolithocholic Acid",
    "Deoxycholic Acid",
    "Isodeoxycholic Acid",
    "Lithocholic Acid",
    "Kynurenic Acid",
    "Kynurenine",
    "Anthranilic Acid",
    "Desaminotyrosine",
    "Niacin",
    "Tyrosine",
    "Tryptamine",
    "Tryptophan",
    "Phenylalanine",
    "Acetate",
    "Butyrate",
    "Propionate",
    "Succinate"
  ),
  molar_mass__gmol = c(
    515.7,
    465.6,
    408.6,
    374.6,
    376.6,
    392.6,
    392.6,
    376.6,
    189.17,
    208.21,
    137.14,
    166.17,
    123.11,
    181.19,
    160.22,
    204.22,
    165.19,
    59.04,
    87.1,
    73.07,
    116.07
  )
)

metab_quant_converted <- metab_quant_anon %>%
  mutate(
    compound = ifelse(compound == "isovaleric-acid", "isovalerate", compound),
    compound = str_to_title(compound)
  ) %>%
  mutate(units = case_when(compound %in% c(
    "Taurocholic Acid",
    "Glycocholic Acid",
    "Cholic Acid",
    "3-Oxolithocholic Acid",
    "Alloisolithocholic Acid",
    "Deoxycholic Acid",
    "Isodeoxycholic Acid",
    "Lithocholic Acid"
  ) ~ "ugmL",
  compound %in% c("Acetate", "Butyrate", "Succinate", "Propionate") ~ "mM",
  TRUE ~ "uM")) %>% 
  right_join(metab_conversions, by = "compound") %>% # right_join to only keep compounds we're interested in
  mutate(mvalue__mM = case_when(units == "ugmL" ~ (mvalue*               #ugmL
                                                  ((1000/                #1000 ml per L
                                                    1000000)/            #1000000 ug per gram
                                                    molar_mass__gmol)*   #gram per mol (molar mass)
                                                    1000                 #1000 mM per 1M                                            
                                                   ),
                                units == "uM" ~ mvalue/                  #uM
                                                1000,                    #1000uM per 1M
                                TRUE ~ mvalue #units are already in mM
                                )
         ) %>% 
  select(sampleID, compound, mvalue__mM)
  

metab_boxplot <- 
  metaphlan_df2 %>% 
  left_join(metaphlan_df_sumry) %>% 
  drop_na(taxid) %>% 
  select(sampleID, diversity_group_abv, db) %>% 
  group_by(sampleID) %>% 
  dplyr::slice(1) %>% 
  left_join(metab_quant_converted) %>% 
  mutate(compound = ifelse(compound == "isovaleric-acid", "isovalerate", compound),
         compound = str_to_title(compound)) %>% 
  ungroup() %>%
  mutate(class = case_when(compound %in% c("Taurocholic Acid", "Glycocholic Acid") ~ "Conjugated Primary Bile Acid",
                           compound %in% c("Cholic Acid") ~ "Primary Bile Acid",
                           compound %in% c("3-Oxolithocholic Acid", "Alloisolithocholic Acid", "Deoxycholic Acid", "Isodeoxycholic Acid",
                                                "Lithocholic Acid") ~ "Secondary Bile Acid",
                           compound %in% c("Threonine", "Glycine", "Tyrosine", "Tyramine", "Serine", "Leucine", "Isoleucine",
                                               "Valine", "Phenylalanine", "Alanine", "Proline", "Aspartate", 
                                               "Methionine", "Glutamate", "Lysine", "Cysteine", "Tryptophan") ~ "Amino Acid",
                           compound %in% c("Acetate", "Butyrate", "Succinate", "Propionate") ~ "Short-Chain Fatty Acid",
                           compound %in% c("Kynurenic Acid", "Anthranilic Acid", "Kynurenine", "Tryptamine") ~ "Kynurenine Metabolite",
                           compound == "Desaminotyrosine" ~ "Phenolic Aromatic",
                           compound == "Niacin" ~ "B-Vitamin",
                              TRUE ~ "Indole"),
         compound = case_when(class == "Conjugated Primary Bile Acid" ~ paste(compound, "(1Conj. BA)"),
                              class == "Primary Bile Acid" ~ paste(compound, "(1 BA)"),
                              class == "Secondary Bile Acid" ~ paste(compound, "(2 BA)"),
                              class == "Short-Chain Fatty Acid" ~ paste(compound, "(SCFA)"),
                              class == "Amino Acid" ~ paste(compound, "(AA)"),
                              class == "Phenolic Aromatic" ~ paste(compound, "(Phen. Arom.)"),
                              class == "Indole" ~ paste0(compound, "(Indole)"),
                              class == "Kynurenine Metabolite" ~ paste(compound, "(Kyn. Metab.)"),
                              class == "B-Vitamin" ~ paste(compound, "(B-Vitamin)")
                              )) %>% 
  drop_na() %>% 
  group_by(compound) %>% 
  mutate(count = length(unique(diversity_group_abv))) %>% 
  filter(count == 4) %>% 
  select(-count) %>%
  mutate(compound = factor(
    compound,
    levels = c(
      "Acetate (SCFA)",                  # 1
      "Propionate (SCFA)",               # 2
      "Butyrate (SCFA)",                 # 3
      "Succinate (SCFA)",                # 4
      "Tyrosine (AA)",                   # 5
      "Tryptophan (AA)",                 # 6
      "Phenylalanine (AA)",              # 7
      "Cholic Acid (1 BA)",             # 8
      "Glycocholic Acid (1Conj. BA)",   # 9
      "Taurocholic Acid (1Conj. BA)",   # 10
      "Deoxycholic Acid (2 BA)",        # 11
      "Lithocholic Acid (2 BA)",        # 12
      "Isodeoxycholic Acid (2 BA)",     # 13
      "3-Oxolithocholic Acid (2 BA)",   # 14
      "Alloisolithocholic Acid (2 BA)", # 15
      "Desaminotyrosine (Phen. Arom.)",  # 16
      "Kynurenine (Kyn. Metab.)",        # 17
      "Anthranilic Acid (Kyn. Metab.)",  # 18
      "Tryptamine (Kyn. Metab.)",        # 19
      "Niacin (B-Vitamin)"               # 20
    )
  ),
  class = factor(
    class,
    levels = c(
      "Short-Chain Fatty Acid",          # 1
      "Amino Acid",                      # 2
      "Primary Bile Acid",               # 3
      "Conjugated Primary Bile Acid",    # 4
      "Secondary Bile Acid",             # 5
      "Indole",                          # 6
      "Phenolic Aromatic",               # 7
      "Kynurenine Metabolite",           # 8
      "B-Vitamin"                        # 9
    )
  )) %>% 
  filter(compound != "Kynurenic Acid") %>% 
  mutate(db = factor(db, levels = c("Liver Transplant", "Healthy Donor")),
         diversity_group_abv = factor(diversity_group_abv, levels = c("Low", "Medium", "High", "Healthy Donor"))) %>% 
  filter(compound %in% c(
    "Acetate (SCFA)",
    "Butyrate (SCFA)",
    "Propionate (SCFA)",
    "Glycocholic Acid (1Conj. BA)",
    "Taurocholic Acid (1Conj. BA)",
    "Cholic Acid (1 BA)",
    "Deoxycholic Acid (2 BA)",
    "Lithocholic Acid (2 BA)",
    "Alloisolithocholic Acid (2 BA)"
    ))


metab_boxplot_stats <-
  metab_boxplot %>% 
  group_by(class, compound) %>% 
  rstatix::wilcox_test(mvalue__mM~diversity_group_abv,
                       comparisons = diversity_comps,
                       p.adjust.method = "none",
                       alternative= "two.sided") %>% 
  rstatix::adjust_pvalue(method = "BH") %>%
  rstatix::add_significance("p.adj") %>% 
  mutate(p.adj = ifelse(p.adj < 0.001, "p.adj < 0.001", paste("p.adj = ", round(p.adj, 2)))) %>% 
  mutate(y.position = case_when(group1 == "High" & group2 == "Healthy Donor" ~ 2.30,
                                group1 == "Medium" & group2 == "High" ~ 2.7,
                                group1 == "Low" & group2 == "High" ~ 3.1,
                                group1 == "Low" & group2 == "Medium" ~ 3.5)) # Set stats brackets to fixed points since the y-scale will be log10 transformed


set.seed(123) # for consistent jittering of points
gg_metab_boxplot <-
ggboxplot(metab_boxplot,
          x = "diversity_group_abv",
          y = "mvalue__mM",
          fill = "db",
          color = "diversity_group_abv",
          alpha = 0.65,
          outlier.shape = NA,
          facet.by = c("class", "compound")) +
    theme(legend.text = et(size = 12, color = "black"),
          legend.title = et(size = 14, color = "black"),
          axis.title.x = eb(),
          axis.title.y = et(size = 12, color = "black"),
          panel.border = eb(),
          strip.background = er(colour="white", fill="white"),
          ) +
    geom_hline(yintercept = 0) + 
    geom_segment(aes(x = 0.35, y = 0, xend = 0.35, yend = Inf)) +
    facet_wrap(~compound, scales = "fixed") +
    stat_pvalue_manual(metab_boxplot_stats, 
                       tip.length = 0.015) +
  geom_point(
    data = metab_boxplot,
    aes(x = diversity_group_abv, y = mvalue__mM, color = diversity_group_abv),
    position = position_jitter(width = 0.2),
    size = 2,
    alpha = 0.65) +
    scale_fill_manual("Cohort", values = rev(pirate_colors)) +
    scale_color_manual("Diversity Group", values = diversity_group_colors) +
    scale_y_log10(limits = c(0.01, 5000),
                  labels = c(0.01, 0.1, 1, 10, 100, 1000),
                  breaks = c(0.01, 0.1, 1, 10, 100, 1000),
                  expand = expansion(mult = c(0.1, 0.2))) +
    ylab("Concentration (mM)\n")

gg_metab_boxplot

cairo_pdf(filename = "./Results/Figure_3.pdf", width = 14, height = 10, onefile = FALSE)

gg_metab_boxplot

invisible(dev.off())
```

### Figure 4A: Enterococcus Expansion and Infection
```{r, figure-4a-metaphlan4-enterococcus-expansion-infection, fig.dim = c(16, 12), message=FALSE, warning=FALSE}
gg_ecoc_all_patients <-
  peri_matrix_all %>%
  distinct(sampleID, .keep_all = T) %>% 
  select(sampleID, patientID, bact_infection_present) %>% 
  mutate(bact_infection_present = ifelse(bact_infection_present == "No", "No Infection", "Infection"),
         bact_infection_present = factor(bact_infection_present, levels = c("Infection", "No Infection"))) %>%
  left_join(metaphlan_peri_anon %>% 
              select(-bact_infection_present) %>% 
              group_by(patientID, sampleID) %>% 
              filter(grepl(x = Species, pattern = "Enterococcus", ignore.case = T)) %>%
              count(sampleID, wt = pctseqs, name = "enterococcus_rel_abundance")) %>% 
  left_join(metaphlan_peri_anon %>% 
              select(patientID, sampleID, Kingdom:pctseqs)) %>% 
  ungroup() %>% 
  arrange(Kingdom, Phylum, Class, Order, Family, Genus) %>%
  mutate(Genus = paste0(Phylum,"-",Order,"-", Family, "-",Genus)) %>% 
  group_by(sampleID) %>%
  arrange(Genus) %>%
  mutate(cum.pct = cumsum(pctseqs),
         y.text = (cum.pct + c(0, cum.pct[-length(cum.pct)]))/2) %>%
  ungroup() %>%
  dplyr::select(-cum.pct) %>%
  mutate(sampleID = ifelse(is.na(sampleID), patientID, sampleID),
         Genus = factor(Genus, levels = unique(Genus)),
         Genus = fct_relevel(Genus, "Firmicutes-Lactobacillales-Enterococcaceae-Enterococcus", after = Inf)
         ) %>% 
  arrange(-enterococcus_rel_abundance) %>%
  group_by(sampleID) %>% 
  ggplot(aes(x=reorder(sampleID, -enterococcus_rel_abundance),y=pctseqs)) +
  geom_bar(aes(fill=Genus),stat="identity") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = et(angle = 90, hjust = 0.5),
        strip.text.x= et(angle=0,size=14),
        strip.background = eb(),
        axis.title.y = et(color = "black", size = 12),
        axis.text.y = et(color = "black", size = 10)) +
  scale_fill_manual(values=metaphlan_pal2) +
  scale_y_continuous(expand = expansion(mult = c(0.005,0.005)))+
  ylab("Shotgun Relative Abundance (%)\n") +
  xlab("")+
  facet_grid(~bact_infection_present, space = "free_x", scales = "free_x")
# gg_ecoc_all_patients


# Discrete heatmap of infections
ecoc_infx_orgs <- 
  peri_criteria_all %>% 
  filter(sampleID %in% peri_matrix_all$sampleID) %>% 
  group_by(patientID, eday) %>%
  arrange(-infx_stool) %>%
  dplyr::slice(1) %>% 
  ungroup() %>% 
  select(patientID, sampleID, bact_infection_present, infx_stool, organism1, micro1.factor) %>%
  distinct() %>% 
  mutate(organism1 = gsub(x = organism1, pattern = "\\s+", replacement = ""),
         organism1 = str_to_lower(string = organism1),
         organism1 = ifelse(grepl(x = organism1, pattern = "enterococcus|enterobacterales|klebsiella|escherichia|citrobacter|proteus|staphyl|clostrid|pseudo|steno|bacteroides|helico"), organism1, "Culture Negative")) %>% 
  group_by(patientID, sampleID, infx_stool) %>%
  mutate(`Enterococcus faecium` = grepl(x = organism1, pattern = "enterococcusfaecium", ignore.case = T),
         `Enterococcus faecalis` = grepl(x = organism1, pattern = "enterococcusfaecalis", ignore.case = T),
         `Enterococcus avium` = grepl(x = organism1, pattern = "enterococcusavium", ignore.case = T),
         `Enterococcus gallinarum` = grepl(x = organism1, pattern = "enterococcusgallinarum", ignore.case = T),
         `Klebsiella pneumoniae` = grepl(x = organism1, pattern = "klebsiellapneumoniae", ignore.case = T),
         `Enterobacter cloaceae` = grepl(x = organism1, pattern = "enterobactercloaceae", ignore.case = T),
         `Escherichia coli` = grepl(x = organism1, pattern = "escherichiacoli", ignore.case = T),
         `Citrobacter freundii` = grepl(x = organism1, pattern = "citrobacterfreundii", ignore.case = T),
         `Proteus mirabilis` = grepl(x = organism1, pattern = "proteusmirabilis", ignore.case = T),
         `Staphylococcus aureus` = grepl(x = organism1, pattern = "staphylococcusaureus", ignore.case = T),
         `Staphylococcus epidermis` = grepl(x = organism1, pattern = "staphylococcusepidermidis", ignore.case = T),
         `Pseudomonas aeruginosa` = grepl(x = organism1, pattern = "pseudomonasaeruginosa", ignore.case = T),
         `Stenotrophmonas maltophilia` = grepl(x = organism1, pattern = "stenotrophmonasmaltophilia", ignore.case = T),
         `Helicobacter pylori` = grepl(x = organism1, pattern = "helicobacterpylori", ignore.case = T),
         `Clostridium difficile` = grepl(x = organism1, pattern = "clostridiumdifficile|clostridioidesdifficile", ignore.case = T),
         `Bacteroides sp.` = grepl(x = organism1, pattern = "bacteroides", ignore.case = T),
         `Culture Negative` = grepl(x = organism1, pattern = "Culture Negative", ignore.case = T)) %>%  
  pivot_longer(-c(patientID:micro1.factor), names_to = "organisms", values_to = "org_presence") %>%
  mutate(organisms = ifelse(bact_infection_present == "No", "No Bacterial Infection", organisms),
         org_presence = ifelse(org_presence == TRUE, 1, 0)) %>% 
  group_by(sampleID, infx_stool, bact_infection_present, organisms) %>% 
  dplyr::slice_max(org_presence) %>% 
  ungroup() %>% 
  filter(org_presence == 1) %>% 
  left_join(metaphlan_peri_anon %>% 
              select(-bact_infection_present) %>% 
              group_by(patientID, sampleID) %>% 
              filter(grepl(x = Species, pattern = "Enterococcus", ignore.case = T)) %>%
              count(sampleID, wt = pctseqs, name = "enterococcus_rel_abundance")) %>% 
  left_join(metaphlan_peri_anon %>% 
              select(patientID, sampleID, Kingdom:pctseqs)) %>% 
  group_by(patientID, sampleID, organisms, org_presence) %>% 
  dplyr::slice(1) %>%
  mutate(sampleID = ifelse(is.na(sampleID), patientID, sampleID)) %>% 
  mutate(org_presence = ifelse(grepl(pattern = "No", x = bact_infection_present, ignore.case = T), 0, org_presence)) %>% 
  ungroup() %>% 
  mutate(organisms = ifelse(bact_infection_present == "Yes" & org_presence == 0, "Other", organisms)) %>% 
  arrange(-enterococcus_rel_abundance) %>% 
  mutate(organisms = ifelse(grepl(x = organisms, pattern = "enterococcus|klebsiella|escherichia|proteus|citrobacter|culture", ignore.case = TRUE), organisms, "Other Bacterial Infection"),
         organisms = ifelse(bact_infection_present == "No", "No Bacterial Infection", organisms))
 
ecoc_infx_orgs_order <-
  ecoc_infx_orgs %>%
  group_by(sampleID, patientID,  organisms) %>% 
  distinct(sampleID, patientID, .keep_all = T) %>% 
  ungroup() %>% 
  mutate(organisms = ifelse(grepl(organisms, pattern = "gallinarum"), "Other Bacterial Infection", organisms),
         org_colors = case_when(grepl(x = organisms, pattern = "enterococcus faecium", ignore.case = T) ~ 1,
                                grepl(x = organisms, pattern = "enterococcus faecalis", ignore.case = T) ~ 2,
                                grepl(x = organisms, pattern = "enterococcus avium", ignore.case = T) ~ 3,
                                grepl(x = organisms, pattern = "klebsiella pneumoniae", ignore.case = T) ~ 4,
                                grepl(x = organisms, pattern = "escherichia coli", ignore.case = T) ~ 5,
                                grepl(x = organisms, pattern = "proteus mirabilis", ignore.case = T) ~ 6,
                                grepl(x = organisms, pattern = "citrobacter freundii", ignore.case = T) ~ 7,
                                grepl(x = organisms, pattern = "other bacterial infection|gallinarum", ignore.case = T) ~ 8,
                                grepl(x = organisms, pattern = "culture negative", ignore.case = T) ~ 9,
                                TRUE ~ 0),
         organisms = as.factor(organisms),
         organisms = factor(organisms, levels = c("Enterococcus faecium",
                                                  "Enterococcus faecalis",
                                                  "Enterococcus avium",
                                                  "Klebsiella pneumoniae",
                                                  "Escherichia coli",
                                                  "Proteus mirabilis",
                                                  "Citrobacter freundii",
                                                  "Other Bacterial Infection",
                                                  "Culture Negative",
                                                  "No Bacterial Infection")),
         org_colors = factor(org_colors, levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "0"))) %>% 
  left_join(ecoc_infx_orgs) %>% 
  mutate(organisms = factor(organisms, levels = c("Enterococcus faecium",
                                                  "Enterococcus faecalis",
                                                  "Enterococcus avium",
                                                  "Klebsiella pneumoniae",
                                                  "Escherichia coli",
                                                  "Proteus mirabilis",
                                                  "Citrobacter freundii",
                                                  "Other Bacterial Infection",
                                                  "Culture Negative",
                                                  "No Bacterial Infection")),
         org_colors = factor(org_colors, levels = c("1", "2", "3", "4", "5","6", "7", "8","9", "0"))) %>%
  drop_na(organisms) %>%
  mutate(bact_infection_present = ifelse(bact_infection_present == "No", "No Infection", "Infection"),
         bact_infection_present = factor(bact_infection_present, levels = c("Infection", "No Infection"))) 

gg_ecoc_infx_orgs <-
 ecoc_infx_orgs_order %>%
  droplevels() %>% 
  ungroup() %>% 
  ggplot(., aes(x = reorder(sampleID, -enterococcus_rel_abundance), y = organisms, fill = as.factor(org_colors)))+
  geom_tile(color = "black") +
  theme_bw()+
  theme(
    axis.title.y = et(color = "black", size = 12),
    axis.text.y = et(color = "black", size = 10),
    axis.ticks.x = eb(),
    axis.text.x = eb(),
    axis.title.x = eb(),
    panel.grid = eb(),
    strip.text = eb()
  ) +
  scale_fill_manual(values = c("#129246",
                               "#0C7A3A",
                               "#08592B",
                               "#FF0000",
                               "#CC0404",
                               "#8A0202",
                               "#5C0202",
                               "#E6C66E",
                               "#BD992D",
                               "#00000000"),
                    labels = c("Enterococcus faecium",
                               "Enterococcus faecalis",
                               "Enterococcus avium",
                               "Klebsiella pneumoniae",
                               "Escherichia coli",
                               "Citrobacter freundii",
                               "Proteus mirabilis",
                               "Other Bacterial Infection",
                               "Culture Negative",
                               "No Bacterial Infection")
                    ) +
  labs(fill = "Infecting Organism",
       y = "Infecting Organism\n")+
  scale_y_discrete(expand = expansion(mult = c(0.005,0.005)),
                   limits = rev(levels(ecoc_infx_orgs_order$organisms))) +
  facet_grid(.~bact_infection_present, space = "free_x", scales = "free_x")

# gg_ecoc_infx_orgs

yingtools2::gg.stack(gg_ecoc_all_patients, 
                     gg_ecoc_infx_orgs,
                     heights = c(1, 0.5),
                     adjust.themes = T)

pdf(file = "./Results/Figure_4A.pdf", height = 8, width = 15, onefile = F)

yingtools2::gg.stack(gg_ecoc_all_patients, 
                     gg_ecoc_infx_orgs,
                     heights = c(1, 0.5),
                     adjust.themes = T)

invisible(dev.off())
```

### Figure 4B: Enterobacterales Expansion and Infection
```{r, figure-4b-metaphlan4-enterobacterales-expansion-infection,fig.dim = c(16, 12), message=FALSE, warning=FALSE}
gg_ebac_all_patients <-
  peri_matrix_all %>%
  distinct(sampleID, .keep_all = T) %>% 
  select(sampleID, patientID, bact_infection_present) %>% 
  mutate(bact_infection_present = ifelse(bact_infection_present == "No", "No Infection", "Infection"),
         bact_infection_present = factor(bact_infection_present, levels = c("Infection", "No Infection"))) %>%
  left_join(metaphlan_peri_anon %>% 
              select(-bact_infection_present) %>% 
              group_by(patientID, sampleID) %>% 
              filter(grepl(x = Species, pattern = "Klebsiella pneumoniae|Escherichia coli|Citrobacter freundii|Proteus mirabilis")) %>%
              count(sampleID, wt = pctseqs, name = "enterobacterales_rel_abundance")) %>% 
  left_join(metaphlan_peri_anon %>% 
              select(patientID, sampleID, Kingdom:pctseqs)) %>% 
  ungroup() %>% 
  arrange(Kingdom, Phylum, Class, Order, Family, Genus) %>%
  mutate(Genus = paste0(Phylum,"-",Order,"-", Family, "-",Genus)) %>% 
  group_by(sampleID) %>%
  arrange(Genus) %>%
  mutate(cum.pct = cumsum(pctseqs),
         y.text = (cum.pct + c(0, cum.pct[-length(cum.pct)]))/2) %>%
  ungroup() %>%
  dplyr::select(-cum.pct) %>%
  mutate(sampleID = ifelse(is.na(sampleID), patientID, sampleID),
         Genus = factor(Genus, levels = unique(Genus)),
         Genus = fct_relevel(Genus, c("Proteobacteria-Enterobacterales-Enterobacteriaceae-Citrobacter",
                                      "Proteobacteria-Enterobacterales-Enterobacteriaceae-Enterobacter",
                                      "Proteobacteria-Enterobacterales-Enterobacteriaceae-Escherichia",
                                      "Proteobacteria-Enterobacterales-Enterobacteriaceae-Klebsiella",
                                      "Proteobacteria-Enterobacterales-Morganellaceae-Proteus"), 
                             after = Inf)) %>% 
  arrange(-enterobacterales_rel_abundance) %>%
  ggplot(aes(x=reorder(sampleID, -enterobacterales_rel_abundance),y=pctseqs)) +
  geom_bar(aes(fill=Genus),stat="identity") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = et(angle = 90, hjust = 0.5),
        strip.text.x= et(angle=0,size=14),
        strip.background = eb(),
        axis.title.y = et(color = "black", size = 12),
        axis.text.y = et(color = "black", size = 10)) +
  scale_fill_manual(values=metaphlan_pal2) +
  scale_y_continuous(expand = expansion(mult = c(0.005,0.005)))+
  ylab("Shotgun Relative Abundance (%)\n") +
  xlab("")+
  facet_grid(~bact_infection_present, space = "free_x", scales = "free_x")
# gg_ebac_all_patients

# Discrete heatmap of infections
ebac_infx_orgs <- 
 peri_criteria_all %>% 
  filter(sampleID %in% peri_matrix_all$sampleID) %>% 
  group_by(patientID, eday) %>%
  arrange(-infx_stool) %>%
  dplyr::slice(1) %>% 
  ungroup() %>% 
  select(patientID, sampleID, bact_infection_present, infx_stool, organism1, micro1.factor) %>%
  distinct() %>% 
  mutate(organism1 = gsub(x = organism1, pattern = "\\s+", replacement = ""),
         organism1 = str_to_lower(string = organism1),
         organism1 = ifelse(grepl(x = organism1, pattern = "enterococcus|enterobacterales|klebsiella|escherichia|citrobacter|proteus|staphyl|clostrid|pseudo|steno|bacteroides|helico"), organism1, "Culture Negative")) %>% 
  group_by(patientID, sampleID, infx_stool) %>%
  mutate(`Enterococcus faecium` = grepl(x = organism1, pattern = "enterococcusfaecium", ignore.case = T),
         `Enterococcus faecalis` = grepl(x = organism1, pattern = "enterococcusfaecalis", ignore.case = T),
         `Enterococcus avium` = grepl(x = organism1, pattern = "enterococcusavium", ignore.case = T),
         `Enterococcus gallinarum` = grepl(x = organism1, pattern = "enterococcusgallinarum", ignore.case = T),
         `Klebsiella pneumoniae` = grepl(x = organism1, pattern = "klebsiellapneumoniae", ignore.case = T),
         `Enterobacter cloaceae` = grepl(x = organism1, pattern = "enterobactercloaceae", ignore.case = T),
         `Escherichia coli` = grepl(x = organism1, pattern = "escherichiacoli", ignore.case = T),
         `Citrobacter freundii` = grepl(x = organism1, pattern = "citrobacterfreundii", ignore.case = T),
         `Proteus mirabilis` = grepl(x = organism1, pattern = "proteusmirabilis", ignore.case = T),
         `Staphylococcus aureus` = grepl(x = organism1, pattern = "staphylococcusaureus", ignore.case = T),
         `Staphylococcus epidermis` = grepl(x = organism1, pattern = "staphylococcusepidermidis", ignore.case = T),
         `Pseudomonas aeruginosa` = grepl(x = organism1, pattern = "pseudomonasaeruginosa", ignore.case = T),
         `Stenotrophmonas maltophilia` = grepl(x = organism1, pattern = "stenotrophmonasmaltophilia", ignore.case = T),
         `Helicobacter pylori` = grepl(x = organism1, pattern = "helicobacterpylori", ignore.case = T),
         `Clostridium difficile` = grepl(x = organism1, pattern = "clostridiumdifficile|clostridioidesdifficile", ignore.case = T),
         `Bacteroides sp.` = grepl(x = organism1, pattern = "bacteroides", ignore.case = T),
         `Culture Negative` = grepl(x = organism1, pattern = "Culture Negative", ignore.case = T)) %>%  
  pivot_longer(-c(patientID:micro1.factor), names_to = "organisms", values_to = "org_presence") %>%
  mutate(organisms = ifelse(bact_infection_present == "No", "No Bacterial Infection", organisms),
         org_presence = ifelse(org_presence == TRUE, 1, 0)) %>% 
  group_by(sampleID, infx_stool, bact_infection_present, organisms) %>% 
  dplyr::slice_max(org_presence) %>% 
  ungroup() %>% 
  filter(org_presence == 1) %>% 
  left_join(metaphlan_peri_anon %>% 
              select(-bact_infection_present) %>% 
              group_by(patientID, sampleID) %>% 
              filter(grepl(x = Species, pattern = "Klebsiella pneumoniae|Escherichia coli|Citrobacter freundii|Proteus mirabilis")) %>%
              count(sampleID, wt = pctseqs, name = "enterobacterales_rel_abundance")) %>% 
  left_join(metaphlan_peri_anon %>% 
              select(patientID, sampleID, Kingdom:pctseqs)) %>% 
  group_by(patientID, sampleID, organisms, org_presence) %>% 
  dplyr::slice(1) %>%
  mutate(sampleID = ifelse(is.na(sampleID), patientID, sampleID)) %>% 
  mutate(org_presence = ifelse(grepl(pattern = "No", x = bact_infection_present, ignore.case = T), 0, org_presence)) %>% 
  ungroup() %>% 
  mutate(organisms = ifelse(bact_infection_present == "Yes" & org_presence == 0, "Other", organisms)) %>% 
  arrange(-enterobacterales_rel_abundance) %>% 
  mutate(organisms = ifelse(grepl(x = organisms, pattern = "enterococcus|klebsiella|escherichia|proteus|citrobacter|culture", ignore.case = TRUE), organisms, "Other Bacterial Infection"),
         organisms = ifelse(bact_infection_present == "No", "No Bacterial Infection", organisms))
  
ebac_infx_orgs_order <-
  ebac_infx_orgs %>%
  group_by(sampleID, patientID,  organisms) %>% 
  distinct(sampleID, patientID, .keep_all = T) %>% 
  ungroup() %>% 
  mutate(organisms = ifelse(grepl(organisms, pattern = "gallinarum"), "Other Bacterial Infection", organisms),
         org_colors = case_when(grepl(x = organisms, pattern = "enterococcus faecium", ignore.case = T) ~ 1,
                                grepl(x = organisms, pattern = "enterococcus faecalis", ignore.case = T) ~ 2,
                                grepl(x = organisms, pattern = "enterococcus avium", ignore.case = T) ~ 3,
                                grepl(x = organisms, pattern = "klebsiella pneumoniae", ignore.case = T) ~ 4,
                                grepl(x = organisms, pattern = "escherichia coli", ignore.case = T) ~ 5,
                                grepl(x = organisms, pattern = "proteus mirabilis", ignore.case = T) ~ 6,
                                grepl(x = organisms, pattern = "citrobacter freundii", ignore.case = T) ~ 7,
                                grepl(x = organisms, pattern = "other bacterial infection|gallinarum", ignore.case = T) ~ 8,
                                grepl(x = organisms, pattern = "culture negative", ignore.case = T) ~ 9,
                                TRUE ~ 0),
         organisms = as.factor(organisms),
         organisms = factor(organisms, levels = c("Klebsiella pneumoniae",
                                                  "Escherichia coli",
                                                  "Proteus mirabilis",
                                                  "Citrobacter freundii",
                                                  "Enterococcus faecium",
                                                  "Enterococcus faecalis",
                                                  "Enterococcus avium",
                                                  "Other Bacterial Infection",
                                                  "Culture Negative",
                                                  "No Bacterial Infection")),
         org_colors = factor(org_colors, levels = c("4", "5", "6", "7", "1", "2", "3", "8", "9", "0"))) %>% 
  left_join(ebac_infx_orgs) %>% 
  mutate(organisms = factor(organisms, levels = c("Klebsiella pneumoniae",
                                                  "Escherichia coli",
                                                  "Proteus mirabilis",
                                                  "Citrobacter freundii",
                                                  "Enterococcus faecium",
                                                  "Enterococcus faecalis",
                                                  "Enterococcus avium",
                                                  "Other Bacterial Infection",
                                                  "Culture Negative",
                                                  "No Bacterial Infection")),
         org_colors = factor(org_colors, levels = c("4", "5", "6", "7", "1", "2", "3", "8", "9", "0"))) %>%
  drop_na(organisms) %>%
  mutate(bact_infection_present = ifelse(bact_infection_present == "No", "No Infection", "Infection"),
         bact_infection_present = factor(bact_infection_present, levels = c("Infection", "No Infection"))) 

gg_ebac_infx_orgs <-
 ebac_infx_orgs_order %>%
  ggplot(., aes(x = reorder(sampleID, -enterobacterales_rel_abundance), y = organisms, fill = as.factor(org_colors)))+
  geom_tile(color = "black") +
  theme_bw()+
  theme(
    axis.title.y = et(color = "black", size = 12),
    axis.text.y = et(color = "black", size = 10),
    axis.ticks.x = eb(),
    axis.text.x = eb(),
    axis.title.x = eb(),
    panel.grid = eb(),
    strip.text = eb()
  ) +
  scale_fill_manual(values = c("#FF0000",
                               "#CC0404",
                               "#8A0202",
                               "#5C0202",
                               "#129246",
                               "#0C7A3A",
                               "#08592B",
                               "#E6C66E",
                               "#BD992D",
                               "#00000000"),
                    labels = c("Klebsiella pneumoniae",
                               "Escherichia coli",
                               "Citrobacter freundii",
                               "Proteus mirabilis",
                               "Enterococcus faecium",
                               "Enterococcus faecalis",
                               "Enterococcus avium",
                               "Other Bacterial Infection",
                               "Culture Negative",
                               "No Bacterial Infection")
                    ) +
  labs(fill = "Infecting Organism",
       y = "Infecting Organism\n")+
  scale_y_discrete(expand = expansion(mult = c(0.005,0.005)),
                   limits = rev(levels(ebac_infx_orgs_order$organisms))) +
  facet_grid(.~bact_infection_present, space = "free_x", scales = "free_x")

# gg_ebac_infx_orgs

yingtools2::gg.stack(gg_ebac_all_patients, 
                     gg_ebac_infx_orgs,
                     heights = c(1, 0.5),
                     adjust.themes = T)

pdf(file = "./Results/Figure_4B.pdf", height = 8, width = 15, onefile = F)

yingtools2::gg.stack(gg_ebac_all_patients, 
                     gg_ebac_infx_orgs,
                     heights = c(1, 0.5),
                     adjust.themes = T)

invisible(dev.off())
```

### Figure 5B: Quant Metabolomics-Enterococcus Expansion
```{r, figure-5b-quant-metabolomics-ecoc-expan, fig.dim = c(16, 8), warning=FALSE, message=FALSE}
# Qual compounds that we can quantify and that are also significant in the volcano analysis for both
signif_compounds <- qual_tot_ecoc_expan %>% 
  rownames_to_column(var = "compound") %>% 
  filter(p.adj <= 0.05 & abs(log2fc_val) > 0.75) %>%
  distinct(compound)

metab_boxplot_ecoc_expan <- 
  peri_matrix_all %>%
  select(sampleID, enterococcus_rel_abundance) %>%
  left_join(sample_lookup) %>% 
  group_by(sampleID) %>%
  slice(1) %>%
  left_join(metab_quant_converted %>% filter(compound %in% signif_compounds$compound)) %>% 
  filter(compound %!in% c("Kynurenine", "Anthranilic Acid")) %>% 
  mutate(compound = ifelse(compound == "isovaleric-acid", "isovalerate", compound),
         compound = str_to_title(compound)) %>% 
  ungroup() %>%
  mutate(db = ifelse(db == "HealthyDonor", "Healthy Donor", "Liver Transplant") %>% 
           factor(., levels = c("Healthy Donor","Liver Transplant")),
         class = case_when(compound %in% c("Taurocholic Acid", "Glycocholic Acid") ~ "Conjugated Primary Bile Acid",
                           compound %in% c("Cholic Acid") ~ "Primary Bile Acid",
                           compound %in% c("3-Oxolithocholic Acid", "Alloisolithocholic Acid", "Deoxycholic Acid", "Isodeoxycholic Acid",
                                                "Lithocholic Acid") ~ "Secondary Bile Acid",
                           compound %in% c("Threonine", "Glycine", "Tyrosine", "Tyramine", "Serine", "Leucine", "Isoleucine",
                                               "Valine", "Phenylalanine", "Alanine", "Proline", "Aspartate", 
                                               "Methionine", "Glutamate", "Lysine", "Cysteine", "Tryptophan") ~ "Amino Acid",
                           compound %in% c("Acetate", "Butyrate", "Succinate", "Propionate") ~ "Short-Chain Fatty Acid",
                           compound %in% c("Kynurenic Acid", "Anthranilic Acid", "Kynurenine", "Tryptamine") ~ "Kynurenine Metabolite",
                           compound == "Desaminotyrosine" ~ "Phenolic Aromatic",
                           compound == "Niacin" ~ "B-Vitamin",
                              TRUE ~ "Indole")) %>% 
  drop_na() %>% 
  group_by(compound) %>% 
  mutate(class = factor(class, levels = c("Conjugated Primary Bile Acid",
                                          "Primary Bile Acid",
                                          "Secondary Bile Acid",
                                          "Short-Chain Fatty Acid",
                                          "Amino Acid",
                                          "Phenolic Aromatic",
                                          "Indole",
                                          "Kynurenine Metabolite",
                                          "B-Vitamin")),
         compound = case_when(class == "Conjugated Primary Bile Acid" ~ paste(compound, "(1Conj. BA)"),
                              class == "Primary Bile Acid" ~ paste(compound, "(1 BA)"),
                              class == "Secondary Bile Acid" ~ paste(compound, "(2 BA)"),
                              class == "Short-Chain Fatty Acid" ~ paste(compound, "(SCFA)"),
                              class == "Amino Acid" ~ paste(compound, "(AA)"),
                              class == "Phenolic Aromatic" ~ paste(compound, "(Phen. Arom.)"),
                              class == "Indole" ~ paste0(compound, "(Indole)"),
                              class == "Kynurenine Metabolite" ~ paste(compound, "(Kyn. Metab.)"),
                              class == "B-Vitamin" ~ paste(compound, "(B-Vitamin)")
                              ),
         compound = factor(compound, levels = c("Acetate (SCFA)", "Butyrate (SCFA)", "Propionate (SCFA)", "Succinate (SCFA)",
           "Taurocholic Acid (1Conj. BA)", "Glycocholic Acid (1Conj. BA)",
                                                "Cholic Acid (1 BA)",
                                                "3-Oxolithocholic Acid (2 BA)", "Alloisolithocholic Acid (2 BA)", "Deoxycholic Acid (2 BA)", "Isodeoxycholic Acid (2 BA)",
                                                "Lithocholic Acid (2 BA)", "Kynurenic Acid (Kyn. Metab.)", "Kynurenine (Kyn. Metab.)", "Anthranilic Acid (Kyn. Metab.)", "Desaminotyrosine", "Niacin",
                                                "Tyrosine", "Tryptamine", "Tryptophan", "Phenylalanine"))
         ) %>% 
  ungroup() %>% 
  drop_na() %>% 
  mutate(enterococcus_expansion = ifelse(enterococcus_rel_abundance >= optimal_cutpoint_rel$optimal_cutpoint[2], 1, 0)) %>% 
  arrange(enterococcus_expansion, sampleID) %>% 
  group_by(compound) %>%
  mutate(enterococcus_expansion_0 = length(mvalue__mM[enterococcus_expansion == "0"]),
         enterococcus_expansion_1 = length(mvalue__mM[enterococcus_expansion == "1"])) %>%
  filter(any(mvalue__mM != 0)) %>% 
  mutate(enterococcus_expansion = ifelse(enterococcus_expansion == "1", "Expansion", "No Expansion")) 
  

metab_boxplot_summary_stats_ecoc_expan <-
  metab_boxplot_ecoc_expan %>% 
  group_by(compound) %>% 
  summarise(y.position = max(mvalue__mM)*1.1)

metab_boxplot_stats_ecoc_expan <-
  metab_boxplot_ecoc_expan %>% 
  group_by(class, compound) %>% 
  rstatix::wilcox_test(mvalue__mM ~ enterococcus_expansion) %>% 
  rstatix::adjust_pvalue(method = "BH") %>%
  rstatix::add_significance("p.adj") %>% 
  mutate(p.adj = ifelse(p.adj < 0.001, "p.adj < 0.001", paste("p.adj = ", round(p.adj, 3)))) %>% 
  add_xy_position(x = "enterococcus_expansion") %>% 
  select(-y.position) %>% 
  left_join(metab_boxplot_summary_stats_ecoc_expan)


gg_metab_boxplot_ecoc_expan <-
ggboxplot(metab_boxplot_ecoc_expan,
          x = "enterococcus_expansion",
          y = "mvalue__mM",
          fill = "enterococcus_expansion",
          alpha = 0.65,
          outlier.shape = NA
          ) +
    theme(legend.text = et(size = 14, color = "black"),
          legend.title = et(size = 14, color = "black"),
          axis.title.x = eb(),
          axis.title.y = et(size = 16, color = "black"),
          strip.text = et(size = 14, color = "black"),
          strip.background = eb(),
          panel.border = eb(),
          axis.line = eb()) +
    facet_wrap(~compound, scales = "free_y", nrow = 2) +
    annotate("segment", x = 0.35, xend = 0.35, y = 0, yend = Inf, colour = "black", linewidth = 1) +
    annotate("segment", x = 0.35, xend = Inf, y = 0, yend = 0, colour = "black") +
    stat_pvalue_manual(metab_boxplot_stats_ecoc_expan, label = "{p.adj}", bracket.size = 0.5) +
    geom_point(data = metab_boxplot_ecoc_expan,
               aes(x = enterococcus_expansion, y = mvalue__mM, fill = enterococcus_expansion),
               position = position_jitter(width = 0.2),
             size = 2, shape = 21, alpha = 0.65, color = "black") +
    scale_fill_manual("Enterococcus", values = c("gray87", "#389458")) +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.2))) +
    ylab("Concentration (mM)\n")

gg_metab_boxplot_ecoc_expan

cairo_pdf(filename = "./Results/Figure_5B.pdf",
         height = 8,
         width = 16,
         onefile = TRUE)

gg_metab_boxplot_ecoc_expan

invisible(dev.off())
```

### Figure 6A: sPLS-DA Diversity Groups and Healthy Donors using Qualitative Metabolomics
```{r, figure-6A-splsda-diversity-groups, fig.dim = c(16, 12), message=FALSE, warning=FALSE}
diversity_metab_mat <-
  metab_qual_anon %>% filter(sampleID %in% first_samps$sampleID) %>% 
  mutate(compound = ifelse(compound == "isovaleric-acid", "isovalerate", compound),
         compound = str_to_title(compound)) %>%
  filter(compound %in% heatmap_cmpds$compound|is.na(compound)) %>%
  ungroup() %>%
  left_join(metaphlan_df_sumry %>% select(sampleID, diversity_group_abv)) %>% 
  group_by(sampleID, compound, diversity_group_abv) %>% 
  summarise(mvalue = mean(mvalue, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate_all(~replace(., is.nan(.), NA)) %>% 
  select(sampleID, compound, mvalue, diversity_group_abv) %>% 
  drop_na(sampleID) %>% 
  pivot_wider(names_from = "compound", values_from = "mvalue") %>% 
  filter(sampleID != "") %>%
  column_to_rownames(var = "sampleID") %>% 
  select(-`NA`, - diversity_group_abv) %>% 
  filter_all(any_vars(!is.na(.)))

diversity_metab_labs <-
  diversity_metab_mat %>% 
  rownames_to_column(var = "sampleID") %>% 
  left_join(metaphlan_df_sumry %>% select(sampleID, diversity_group_abv)) %>% 
  droplevels() %>% 
  pull(diversity_group_abv)

dim(diversity_metab_mat) #123 93 (means 93 compounds and 102 LT patients/infections)
length(diversity_metab_labs) #123 (means 102 LT patients/infections)

# Begin model training
set.seed(1234)
diversity_train <- sample(1:nrow(diversity_metab_mat), as.integer(0.7*nrow(diversity_metab_mat))) # randomly select 70% samples in training
diversity_test <- setdiff(1:nrow(diversity_metab_mat), diversity_train) # rest is part of the test set

# store matrices into training and test set:
diversity_metab_mat.train <- diversity_metab_mat[diversity_train, ]
diversity_metab_mat.test <- diversity_metab_mat[diversity_test,]
diversity_metab_labs.train <- diversity_metab_labs[diversity_train]
diversity_metab_labs.test <- diversity_metab_labs[diversity_test]

# Train the model to tune hyperparameters
# Initial model to find optimal number of components to include
set.seed(1234)
diversity_train_splsda <- mixOmics::splsda(diversity_metab_mat.train, diversity_metab_labs.train, ncomp = 5)

# Performance assessment
## 5-fold, 50-repeat cross validation
set.seed(1234)
diversity_train_plsda_perf <-
  perf(
    diversity_train_splsda,
    validation = "Mfold",
    folds = 5,
    progressBar = FALSE,
    auc = TRUE,
    nrepeat = 50
  )

plot(
  diversity_train_plsda_perf,
  col = color.mixo(5:7),
  sd = FALSE,
  auc = TRUE,
  legend.position = "horizontal"
) # ncomp = 4 might be best for classification error rate and max.dist

# Number of optimal variables to select for each component
diversity_train_keepX <- c(1:10,  seq(20, 130, 10))

set.seed(123)
diversity_train_tune_splsda <-
  mixOmics::tune.splsda(
    diversity_metab_mat.train,
    diversity_metab_labs.train,
    ncomp = 4, # Choose 4 components (max) to be safe
    validation = 'Mfold',
    folds = 5,
    dist = 'max.dist',
    progressBar = FALSE,
    auc = TRUE,
    measure = "BER",
    test.keepX = diversity_train_keepX,
    nrepeat = 50
  )

plot(diversity_train_tune_splsda, col = color.jet(4))

diversity_train_error <- diversity_train_tune_splsda$error.rate
diversity_train_ncomp <- diversity_train_tune_splsda$choice.ncomp$ncomp # optimal number of components based on t-tests on the error rate
diversity_train_ncomp #2 components are optimal

diversity_train_select_keepX <- diversity_train_tune_splsda$choice.keepX[1:diversity_train_ncomp]  # optimal number of variables to select per component
diversity_train_select_keepX


# Final Model
diversity_train_splsda_final <-
  mixOmics::splsda(diversity_metab_mat.train, diversity_metab_labs.train, ncomp = diversity_train_ncomp, keepX = diversity_train_select_keepX)

# Test the model
diversity_predict_train_splsda_final <- predict(diversity_train_splsda_final, diversity_metab_mat.test, 
                                dist = "max.dist")

diversity_predict_train_comp2 <- diversity_predict_train_splsda_final$class$max.dist[,diversity_train_ncomp]

diversity_union <- union(diversity_predict_train_comp2, diversity_metab_labs.test)

confusionMatrix(table(factor(diversity_predict_train_comp2, diversity_union,
                             levels = c("Low", "Medium", "High")),
                      factor(diversity_metab_labs.test, diversity_union,
                             levels = c("Low", "Medium", "High"))))

diversity_train_background <- background.predict(diversity_train_splsda_final,
                                            comp.predicted = 2,
                                            xlim = c(-20,20),
                                            ylim = c(-20,20),
                                            dist = "centroids.dist") 

# Model metrics for all samples
diversity_tot <- predict(diversity_train_splsda_final,
                        diversity_metab_mat,
                        dist = "max.dist")

diversity_tot_predict <- diversity_tot$class$max.dist[,diversity_train_ncomp]

diversity_tot_union <- union(diversity_tot_predict, diversity_metab_labs)

diversity_cm <- confusionMatrix(table(factor(diversity_tot_predict, diversity_tot_union,
                                        levels = c("Low", "Medium", "High")),
                      factor(diversity_metab_labs, diversity_tot_union,
                             levels = c("Low", "Medium", "High"))),
                        )
diversity_cm

# Additional model measures
diversity_epi <- mltest::ml_test(predicted = factor(diversity_tot_predict, levels = c("Low", "Medium", "High")),
                                 true = factor(diversity_metab_labs, levels = c("Low", "Medium", "High")))

diversity_cm_names <- diversity_cm$table
colnames(diversity_cm_names) <- c("Actual\nLow", "Actual\nMedium", "Actual\nHigh")
rownames(diversity_cm_names) <- c("Predicted\nLow", "Predicted\nMedium", "Predicted\nHigh")

diversity_confusion_df <- diversity_cm_names %>% 
  t() 

{
pdf(file = "./Results/Figure_6A.pdf", height = 10, width = 10)
plotIndiv(
  diversity_train_splsda_final,
  xlim = c(min(diversity_train_splsda_final$variates$X[,1])*1.05, max(diversity_train_splsda_final$variates$X[,1])*1.8),
  ylim = c(min(diversity_train_splsda_final$variates$X[,2])*1.15, max(diversity_train_splsda_final$variates$X[,2])*1.8),
  comp = c(1,2),
  pch = 1,
  ind.names = FALSE,
  legend = FALSE,
  background = diversity_train_background,
  col = c("#3A001E", "#8A0246", "#C20463"),
  star = TRUE,
  point.lwd = 0.5,
  title = NULL,
  size.title = 0.00001,
  style = "graphics",
  legend.title = "Diversity Group",
  X.label = paste0("Component 1 (", round(diversity_train_splsda_final$prop_expl_var$X[1] * 100), "%)"),
  Y.label = paste0("Component 2 (", round(diversity_train_splsda_final$prop_expl_var$X[2] * 100), "%)")
) 
addtable2plot(
  y = -6.68,
  x = min(diversity_train_splsda_final$variates$X[,1])*1.1,
  diversity_confusion_df,
  bty = "o",
  display.rownames = TRUE,
  hlines = TRUE,
  vlines = TRUE,
  cex = 0.75,
  bg = "white"
)
text(
  y = -4.75,
  x = min(diversity_train_splsda_final$variates$X[,1])*1.1,
  paste0("Overall ACC = ", 
                    paste0(formatC(round(diversity_cm$overall[1], 3)*100, digits = 1, format = "f"), "%"), " [",
                    paste0(formatC(round(diversity_cm$overall[3], 3)*100, digits = 1, format = "f"), "%"),", ",
                    paste0(formatC(round(diversity_cm$overall[4], 3)*100, digits = 1, format = "f"), "%"),"]"),
    cex = 0.75, adj = 0)

text(
  y = 3,
  x = 1.5,
  paste0("Low Diversity ACC = ", round(diversity_epi$balanced.accuracy[1]*100, 1), "%"),
    cex = 0.75, adj = 0, col = "#3A001E")
text(
  y = 2.7,
  x = 1.5,
  paste0("Low Diversity Sens. = ", round(diversity_epi$precision[1]*100, 1), "%"),
    cex = 0.75, adj = 0, col = "#3A001E")
text(
  y = 2.4,
  x = 1.5,
  paste0("Low Diversity Spec. = ", round(diversity_epi$specificity[1]*100, 1), "%"),
    cex = 0.75, adj = 0, col = "#3A001E")
# text(
#   y = 2.1,
#   x = 1.5,
#   paste0("Low Diversity OR = ", round(diversity_epi$DOR[1], 1)),
#     cex = 0.75, adj = 0, col = "#3A001E")

text(
  y = -5,
  x = 1.5,
  paste0("Medium Diversity ACC = ", round(diversity_epi$balanced.accuracy[2]*100, 1), "%"),
    cex = 0.75, adj = 0, col = "#8A0246")
text(
  y = -5.3,
  x = 1.5,
  paste0("Medium Diversity Sens. = ", round(diversity_epi$precision[2]*100, 1), "%"),
    cex = 0.75, adj = 0, col = "#8A0246")
text(
  y = -5.6,
  x = 1.5,
  paste0("Medium Diversity Spec. = ", round(diversity_epi$specificity[2]*100, 1), "%"),
    cex = 0.75, adj = 0, col = "#8A0246")
# text(
#   y = -5.9,
#   x = 1.5,
#   paste0("Medium Diversity OR = ", round(diversity_epi$DOR[2], 1)),
#     cex = 0.75, adj = 0, col = "#8A0246")

text(
  y = 3,
  x = min(diversity_train_splsda_final$variates$X[,1])*1.05,
  paste0("High Diversity ACC = ", round(diversity_epi$balanced.accuracy[3]*100, 1), "%"),
    cex = 0.75, adj = 0, col = "#C20463")
text(
  y = 2.7,
  x = min(diversity_train_splsda_final$variates$X[,1])*1.05,
  paste0("High Diversity Sens. = ", round(diversity_epi$precision[3]*100, 1), "%"),
    cex = 0.75, adj = 0, col = "#C20463")
text(
  y = 2.4,
  x = min(diversity_train_splsda_final$variates$X[,1])*1.05,
  paste0("High Diversity Spec. = ", round(diversity_epi$specificity[3]*100, 1), "%"),
    cex = 0.75, adj = 0, col = "#C20463")
# text(
#   y = 2.1,
#   x = min(diversity_train_splsda_final$variates$X[,1])*1.05,
#   paste0("High Diversity OR = ", round(diversity_epi$DOR[3], 1)),
#     cex = 0.75, adj = 0, col = "#C20463")

invisible(dev.off())
}

plotIndiv(
  diversity_train_splsda_final,
  xlim = c(min(diversity_train_splsda_final$variates$X[,1])*1.05, max(diversity_train_splsda_final$variates$X[,1])*1.8),
  ylim = c(min(diversity_train_splsda_final$variates$X[,2])*1.15, max(diversity_train_splsda_final$variates$X[,2])*1.8),
  comp = c(1,2),
  pch = 1,
  ind.names = FALSE,
  legend = FALSE,
  background = diversity_train_background,
  col = c("#3A001E", "#8A0246", "#C20463"),
  star = TRUE,
  point.lwd = 0.5,
  title = NULL,
  size.title = 0.00001,
  style = "graphics",
  legend.title = "Diversity Group",
  X.label = paste0("Component 1 (", round(diversity_train_splsda_final$prop_expl_var$X[1] * 100), "%)"),
  Y.label = paste0("Component 2 (", round(diversity_train_splsda_final$prop_expl_var$X[2] * 100), "%)")
) 
addtable2plot(
  y = -6.68,
  x = min(diversity_train_splsda_final$variates$X[,1])*1.1,
  diversity_confusion_df,
  bty = "o",
  display.rownames = TRUE,
  hlines = TRUE,
  vlines = TRUE,
  cex = 0.75,
  bg = "white"
)
text(
  y = -4.75,
  x = min(diversity_train_splsda_final$variates$X[,1])*1.1,
  paste0("Overall ACC = ", 
                    paste0(formatC(round(diversity_cm$overall[1], 3)*100, digits = 1, format = "f"), "%"), " [",
                    paste0(formatC(round(diversity_cm$overall[3], 3)*100, digits = 1, format = "f"), "%"),", ",
                    paste0(formatC(round(diversity_cm$overall[4], 3)*100, digits = 1, format = "f"), "%"),"]"),
    cex = 0.75, adj = 0)

text(
  y = 3,
  x = 1.5,
  paste0("Low Diversity ACC = ", round(diversity_epi$balanced.accuracy[1]*100, 1), "%"),
    cex = 0.75, adj = 0, col = "#3A001E")
text(
  y = 2.7,
  x = 1.5,
  paste0("Low Diversity Sens. = ", round(diversity_epi$precision[1]*100, 1), "%"),
    cex = 0.75, adj = 0, col = "#3A001E")
text(
  y = 2.4,
  x = 1.5,
  paste0("Low Diversity Spec. = ", round(diversity_epi$specificity[1]*100, 1), "%"),
    cex = 0.75, adj = 0, col = "#3A001E")
# text(
#   y = 2.1,
#   x = 1.5,
#   paste0("Low Diversity OR = ", round(diversity_epi$DOR[1], 1)),
#     cex = 0.75, adj = 0, col = "#3A001E")

text(
  y = -5,
  x = 1.5,
  paste0("Medium Diversity ACC = ", round(diversity_epi$balanced.accuracy[2]*100, 1), "%"),
    cex = 0.75, adj = 0, col = "#8A0246")
text(
  y = -5.3,
  x = 1.5,
  paste0("Medium Diversity Sens. = ", round(diversity_epi$precision[2]*100, 1), "%"),
    cex = 0.75, adj = 0, col = "#8A0246")
text(
  y = -5.6,
  x = 1.5,
  paste0("Medium Diversity Spec. = ", round(diversity_epi$specificity[2]*100, 1), "%"),
    cex = 0.75, adj = 0, col = "#8A0246")
# text(
#   y = -5.9,
#   x = 1.5,
#   paste0("Medium Diversity OR = ", round(diversity_epi$DOR[2], 1)),
#     cex = 0.75, adj = 0, col = "#8A0246")

text(
  y = 3,
  x = min(diversity_train_splsda_final$variates$X[,1])*1.05,
  paste0("High Diversity ACC = ", round(diversity_epi$balanced.accuracy[3]*100, 1), "%"),
    cex = 0.75, adj = 0, col = "#C20463")
text(
  y = 2.7,
  x = min(diversity_train_splsda_final$variates$X[,1])*1.05,
  paste0("High Diversity Sens. = ", round(diversity_epi$precision[3]*100, 1), "%"),
    cex = 0.75, adj = 0, col = "#C20463")
text(
  y = 2.4,
  x = min(diversity_train_splsda_final$variates$X[,1])*1.05,
  paste0("High Diversity Spec. = ", round(diversity_epi$specificity[3]*100, 1), "%"),
    cex = 0.75, adj = 0, col = "#C20463")
# text(
#   y = 2.1,
#   x = min(diversity_train_splsda_final$variates$X[,1])*1.05,
#   paste0("High Diversity OR = ", round(diversity_epi$DOR[3], 1)),
#     cex = 0.75, adj = 0, col = "#C20463")
```

### Figure 6B: sPLS-DA Any Bacterial Infection using Qualitative Metabolomics
```{r, figure-6B-splsda-any-infection, fig.dim = c(16, 12), message=FALSE, warning=FALSE}
infx_metab_mat <-
  metab_qual_anon %>%
  mutate(compound = ifelse(compound == "isovaleric-acid", "isovalerate", compound),
       compound = str_to_title(compound)) %>%
  filter(compound %in% heatmap_cmpds$compound|is.na(compound)) %>%
  ungroup() %>%
  right_join(peri_matrix_all %>% select(sampleID, bact_infection_present)) %>% 
  group_by(sampleID, compound, bact_infection_present) %>% 
  summarise(mvalue = mean(mvalue, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate_all(~replace(., is.nan(.), NA)) %>% 
  select(sampleID, compound, mvalue, bact_infection_present) %>% 
  drop_na(sampleID) %>% 
  pivot_wider(names_from = "compound", values_from = "mvalue") %>% 
  filter(sampleID != "") %>%
  mutate(bact_infection_present = ifelse(grepl(x = bact_infection_present, pattern = "No"), "No Infection", "Infection")) %>% 
  select(-bact_infection_present) %>% 
  column_to_rownames(var = "sampleID") %>% 
  select(-`NA`) %>% 
  filter_all(any_vars(!is.na(.)))

  
infx_metab_labs <-
  infx_metab_mat %>% 
  rownames_to_column(var = "sampleID") %>% 
  left_join(peri_matrix_all %>% select(sampleID, bact_infection_present)) %>% 
  mutate(bact_infection_present = ifelse(grepl(x = bact_infection_present, pattern = "No"), "No Infection", "Infection")) %>% 
  pull(bact_infection_present)

dim(infx_metab_mat) #108 93 (means 93 compounds and 108 LT patients/infections)
length(infx_metab_labs) #108 (means 108 LT patients/infections)

# Begin model training
set.seed(1234)
infx_train <- sample(1:nrow(infx_metab_mat), as.integer(0.7*nrow(infx_metab_mat))) # randomly select 70% samples in training
infx_test <- setdiff(1:nrow(infx_metab_mat), infx_train) # rest is part of the test set

# store matrices into training and test set:
infx_metab_mat.train <- infx_metab_mat[infx_train, ]
infx_metab_mat.test <- infx_metab_mat[infx_test,]
infx_metab_labs.train <- infx_metab_labs[infx_train]
infx_metab_labs.test <- infx_metab_labs[infx_test]

# Train the model to tune hyperparameters
# Initial model to find optimal number of components to include
set.seed(1234)
infx_train_splsda <- mixOmics::splsda(infx_metab_mat.train, infx_metab_labs.train, ncomp = 5)

# Performance assessment
## 5-fold, 100-repeat cross validation
set.seed(1234)
infx_train_plsda_perf <-
  perf(
    infx_train_splsda,
    validation = "Mfold",
    folds = 5,
    progressBar = FALSE,
    auc = TRUE,
    nrepeat = 50
  )

plot(
  infx_train_plsda_perf,
  col = color.mixo(5:7),
  sd = FALSE,
  auc = TRUE,
  legend.position = "horizontal"
) # ncomp = 1 is best for classification error rate and max.dist

# Number of optimal variables to select for each component
infx_train_keepX <- c(1:10,  seq(20, 108, 10))

set.seed(123)
infx_train_tune_splsda <-
  mixOmics::tune.splsda(
    infx_metab_mat.train,
    infx_metab_labs.train,
    ncomp = 3, # Choose 3 components (max) to be safe
    validation = 'Mfold',
    folds = 5,
    dist = 'max.dist',
    progressBar = FALSE,
    auc = TRUE,
    measure = "BER",
    test.keepX = infx_train_keepX,
    nrepeat = 50
  )

plot(infx_train_tune_splsda, col = color.jet(3))

infx_train_error <- infx_train_tune_splsda$error.rate
infx_train_ncomp <- infx_train_tune_splsda$choice.ncomp$ncomp # optimal number of components based on t-tests on the error rate
infx_train_ncomp #1 component is optimal

infx_train_select_keepX <- infx_train_tune_splsda$choice.keepX[1:(infx_train_ncomp + 1)]  # optimal number of variables to select per component
infx_train_select_keepX


# Final Model
infx_train_splsda_final <-
  mixOmics::splsda(infx_metab_mat.train, infx_metab_labs.train, ncomp = (infx_train_ncomp + 1), keepX = infx_train_select_keepX)

# Test the model
infx_predict_train_splsda_final <- predict(infx_train_splsda_final, infx_metab_mat.test, 
                                dist = "max.dist")

infx_predict_train_comp2 <- infx_predict_train_splsda_final$class$max.dist[,(infx_train_ncomp + 1)]

infx_union <- union(infx_predict_train_comp2, infx_metab_labs.test)

confusionMatrix(table(factor(infx_predict_train_comp2, infx_union),
                      factor(infx_metab_labs.test, infx_union)),
                 positive = "Infection")

infx_train_background <- background.predict(infx_train_splsda_final,
                                            comp.predicted = 2,
                                            xlim = c(-20,20),
                                            ylim = c(-20,20),
                                            dist = "centroids.dist") 

# Model metrics for all samples
infx_tot <- predict(infx_train_splsda_final,
                        infx_metab_mat,
                        dist = "max.dist")

infx_tot_predict <- infx_tot$class$max.dist[,(infx_train_ncomp + 1)]

infx_tot_union <- union(infx_tot_predict, infx_metab_labs)

infx_cm <- confusionMatrix(table(factor(infx_tot_predict, infx_tot_union,
                                        levels = c("Infection", "No Infection")),
                      factor(infx_metab_labs, infx_tot_union,
                             levels = c("Infection", "No Infection"))),
                      positive = "Infection")
infx_cm

# Additional model measures
infx_epi <- epiR::epi.tests(table(infx_tot_predict, infx_metab_labs), conf.level = 0.95)

infx_confusion_df <- infx_epi$tab %>% 
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "actual") %>% 
  mutate(actual = case_when(grepl(actual, pattern = "+", fixed = TRUE) ~ "Actual\nInfection",
                            grepl(actual, pattern = "-", fixed = TRUE) ~ "Actual\nNo Infection",
                            TRUE ~ "Total")) %>% 
  dplyr::rename("Predicted\nInfection" = "Test +",
                "Predicted\nNo Infection" = "Test -") %>% 
  column_to_rownames(var = "actual")

{
pdf(file = "./Results/Figure_6B.pdf", height = 10, width = 10)
plotIndiv(
  infx_train_splsda_final,
  comp = c(1,2),
  pch = 1,
  ind.names = FALSE,
  legend = FALSE,
  background = infx_train_background,
  col = c("goldenrod", "gray75"),
  star = TRUE,
  point.lwd = 0.5,
  title = NULL,
  size.title = 0.00001,
  style = "graphics",
  legend.title = "Infection Group",
  X.label = paste0("Component 1 (", round(infx_train_splsda_final$prop_expl_var$X[1] * 100), "%)"),
  Y.label = paste0("Component 2 (", round(infx_train_splsda_final$prop_expl_var$X[2] * 100), "%)")
) 
addtable2plot(
  y = max(infx_train_splsda_final$variates$X[,2])*0.1,
  x = min(infx_train_splsda_final$variates$X[,1]),
  infx_confusion_df,
  bty = "o",
  display.rownames = TRUE,
  hlines = TRUE,
  vlines = TRUE,
  cex = 0.75,
  bg = "white"
)
text(
  y = -4,
  x = -2.75,
  substitute(bold("Infection")), cex = 1.75, adj = 0, col = "goldenrod")
text(
  y = 4,
  x = -1,
  substitute(bold("No Infection")), cex = 1.75, adj = 0, col = "gray70")
text(
  y = max(infx_train_splsda_final$variates$X[,2])*1,
  x = min(infx_train_splsda_final$variates$X[,1]),
  paste0("ACC = ", 
                    round(infx_cm$overall[1]*100, 1), "% [",
                    round(infx_cm$overall[3]*100, 1), "%, ",
                    round(infx_cm$overall[4]*100, 1), "%]"),
    cex = 0.75, adj = 0)
text(
  y = max(infx_train_splsda_final$variates$X[,2])*0.9,
  x = min(infx_train_splsda_final$variates$X[,1]),
  paste0("Sens. = ", 
                    paste0(formatC(round(infx_epi$detail[2][3,], 3)*100, digits = 1, format = "f"), "%"), " [",
                    paste0(formatC(round(infx_epi$detail[3][3,], 3)*100, digits = 1, format = "f"), "%"),", ",
                    paste0(formatC(round(infx_epi$detail[4][3,], 3)*100, digits = 1, format = "f"), "%"),"]"),
    cex = 0.75, adj = 0)
text(
  y = max(infx_train_splsda_final$variates$X[,2])*0.8,
  x = min(infx_train_splsda_final$variates$X[,1]),
  paste0("Spec. = ", 
                    paste0(formatC(round(infx_epi$detail[2][4,], 3)*100, digits = 1, format = "f"), "%"), " [",
                    paste0(formatC(round(infx_epi$detail[3][4,], 3)*100, digits = 1, format = "f"), "%"),", ",
                    paste0(formatC(round(infx_epi$detail[4][4,], 3)*100, digits = 1, format = "f"), "%"),"]"),
    cex = 0.75, adj = 0)
text(
  y = max(infx_train_splsda_final$variates$X[,2])*0.7,
  x = min(infx_train_splsda_final$variates$X[,1]),
  paste0("OR = ",
                    formatC(round(infx_epi$detail[2][6,], 3), digits = 1, format = "f"),  " [",
                    formatC(round(infx_epi$detail[3][6,], 3), digits = 1, format = "f"), ", ",
                    formatC(round(infx_epi$detail[4][6,], 3), digits = 1, format = "f"), "]"),
    cex = 0.75, adj = 0)
invisible(dev.off())
}

 plotIndiv(
  infx_train_splsda_final,
  comp = c(1,2),
  pch = 1,
  ind.names = FALSE,
  legend = FALSE,
  background = infx_train_background,
  col = c("goldenrod", "gray87"),
  star = TRUE,
  point.lwd = 0.5,
  title = NULL,
  size.title = 0.00001,
  style = "graphics",
  legend.title = "Infection Group",
  X.label = paste0("Component 1 (", round(infx_train_splsda_final$prop_expl_var$X[1] * 100), "%)"),
  Y.label = paste0("Component 2 (", round(infx_train_splsda_final$prop_expl_var$X[2] * 100), "%)")
) 
addtable2plot(
  y = max(infx_train_splsda_final$variates$X[,2])*0.1,
  x = min(infx_train_splsda_final$variates$X[,1]),
  infx_confusion_df,
  bty = "o",
  display.rownames = TRUE,
  hlines = TRUE,
  vlines = TRUE,
  cex = 0.75,
  bg = "white"
)
text(
  y = -4,
  x = -2.75,
  substitute(bold("Infection")), cex = 1.75, adj = 0, col = "goldenrod")
text(
  y = 4,
  x = -1,
  substitute(bold("No Infection")), cex = 1.75, adj = 0, col = "gray70")
text(
  y = max(infx_train_splsda_final$variates$X[,2])*1,
  x = min(infx_train_splsda_final$variates$X[,1]),
  paste0("ACC = ", 
                    round(infx_cm$overall[1]*100, 1), "% [",
                    round(infx_cm$overall[3]*100, 1), "%, ",
                    round(infx_cm$overall[4]*100, 1), "%]"),
    cex = 0.75, adj = 0)
text(
  y = max(infx_train_splsda_final$variates$X[,2])*0.9,
  x = min(infx_train_splsda_final$variates$X[,1]),
  paste0("Sens. = ", 
                    paste0(formatC(round(infx_epi$detail[2][3,], 3)*100, digits = 1, format = "f"), "%"), " [",
                    paste0(formatC(round(infx_epi$detail[3][3,], 3)*100, digits = 1, format = "f"), "%"),", ",
                    paste0(formatC(round(infx_epi$detail[4][3,], 3)*100, digits = 1, format = "f"), "%"),"]"),
    cex = 0.75, adj = 0)
text(
  y = max(infx_train_splsda_final$variates$X[,2])*0.8,
  x = min(infx_train_splsda_final$variates$X[,1]),
  paste0("Spec. = ", 
                    paste0(formatC(round(infx_epi$detail[2][4,], 3)*100, digits = 1, format = "f"), "%"), " [",
                    paste0(formatC(round(infx_epi$detail[3][4,], 3)*100, digits = 1, format = "f"), "%"),", ",
                    paste0(formatC(round(infx_epi$detail[4][4,], 3)*100, digits = 1, format = "f"), "%"),"]"),
    cex = 0.75, adj = 0)
text(
  y = max(infx_train_splsda_final$variates$X[,2])*0.7,
  x = min(infx_train_splsda_final$variates$X[,1]),
  paste0("OR = ",
                    formatC(round(infx_epi$detail[2][6,], 3), digits = 1, format = "f"),  " [",
                    formatC(round(infx_epi$detail[3][6,], 3), digits = 1, format = "f"), ", ",
                    formatC(round(infx_epi$detail[4][6,], 3), digits = 1, format = "f"), "]"),
    cex = 0.75, adj = 0)
```

### Combine Figure 6A + 6B: sPLS-DA Diversity Groups and Any Bacterial Infection 
```{r, figure-6-combined, fig.dim = c(20, 10), message=FALSE, warning=FALSE}
{
pdf(file = "./Results/Figure_6.pdf", height = 10, width = 18)  
par(mfrow=c(1,2))

# Figure 6A
plotIndiv(
  diversity_train_splsda_final,
  xlim = c(min(diversity_train_splsda_final$variates$X[,1])*1.05, max(diversity_train_splsda_final$variates$X[,1])*1.8),
  ylim = c(min(diversity_train_splsda_final$variates$X[,2])*1.15, max(diversity_train_splsda_final$variates$X[,2])*1.8),
  comp = c(1,2),
  pch = 1,
  ind.names = FALSE,
  legend = FALSE,
  background = diversity_train_background,
  col = c("#3A001E", "#8A0246", "#C20463"),
  star = TRUE,
  point.lwd = 0.5,
  title = NULL,
  size.title = 0.00001,
  style = "graphics",
  legend.title = "Diversity Group",
  X.label = paste0("Component 1 (", round(diversity_train_splsda_final$prop_expl_var$X[1] * 100), "%)"),
  Y.label = paste0("Component 2 (", round(diversity_train_splsda_final$prop_expl_var$X[2] * 100), "%)")
) 
# addtable2plot(
#   y = -4.68,
#   x = min(diversity_train_splsda_final$variates$X[,1])*1.1,
#   diversity_confusion_df,
#   bty = "o",
#   display.rownames = TRUE,
#   hlines = TRUE,
#   vlines = TRUE,
#   cex = 1.35,
#   bg = "white"
# )
text(
  y = -6.5,
  x = min(diversity_train_splsda_final$variates$X[,1])*1.1,
  paste0("Overall ACC = ", 
                    paste0(formatC(round(diversity_cm$overall[1], 3)*100, digits = 1, format = "f"), "%"), " [",
                    paste0(formatC(round(diversity_cm$overall[3], 3)*100, digits = 1, format = "f"), "%"),", ",
                    paste0(formatC(round(diversity_cm$overall[4], 3)*100, digits = 1, format = "f"), "%"),"]"),
    cex = 1.35, adj = 0)

text(
  y = 3,
  x = 0.5,
  paste0("Low Diversity ACC = ", round(diversity_epi$balanced.accuracy[1]*100, 1), "%"),
    cex = 1.35, adj = 0, col = "#3A001E")
text(
  y = 2.7,
  x = 0.5,
  paste0("Low Diversity Sens. = ", round(diversity_epi$precision[1]*100, 1), "%"),
    cex = 1.35, adj = 0, col = "#3A001E")
text(
  y = 2.4,
  x = 0.5,
  paste0("Low Diversity Spec. = ", round(diversity_epi$specificity[1]*100, 1), "%"),
    cex = 1.35, adj = 0, col = "#3A001E")
# text(
#   y = 2.1,
#   x = 0.5,
#   paste0("Low Diversity OR = ", round(diversity_epi$DOR[1], 1)),
#     cex = 1.35, adj = 0, col = "#3A001E")

text(
  y = -5.5,
  x = 0,
  paste0("Medium Diversity ACC = ", round(diversity_epi$balanced.accuracy[2]*100, 1), "%"),
    cex = 1.35, adj = 0, col = "#8A0246")
text(
  y = -5.8,
  x = 0,
  paste0("Medium Diversity Sens. = ", round(diversity_epi$precision[2]*100, 1), "%"),
    cex = 1.35, adj = 0, col = "#8A0246")
text(
  y = -6.1,
  x = 0,
  paste0("Medium Diversity Spec. = ", round(diversity_epi$specificity[2]*100, 1), "%"),
    cex = 1.35, adj = 0, col = "#8A0246")
# text(
#   y = -7.4,
#   x = 0,
#   paste0("Medium Diversity OR = ", round(diversity_epi$DOR[2], 1)),
#     cex = 1.05, adj = 0, col = "#8A0246")

text(
  y = 3,
  x = min(diversity_train_splsda_final$variates$X[,1])*1.05,
  paste0("High Diversity ACC = ", round(diversity_epi$balanced.accuracy[3]*100, 1), "%"),
    cex = 1.35, adj = 0, col = "#C20463")
text(
  y = 2.7,
  x = min(diversity_train_splsda_final$variates$X[,1])*1.05,
  paste0("High Diversity Sens. = ", round(diversity_epi$precision[3]*100, 1), "%"),
    cex = 1.35, adj = 0, col = "#C20463")
text(
  y = 2.4,
  x = min(diversity_train_splsda_final$variates$X[,1])*1.05,
  paste0("High Diversity Spec. = ", round(diversity_epi$specificity[3]*100, 1), "%"),
    cex = 1.35, adj = 0, col = "#C20463")
# text(
#   y = 2.1,
#   x = min(diversity_train_splsda_final$variates$X[,1])*1.05,
#   paste0("High Diversity OR = ", round(diversity_epi$DOR[3], 1)),
#     cex = 1.35, adj = 0, col = "#C20463")

# Figure 6B
plotIndiv(
  infx_train_splsda_final,
  comp = c(1,2),
  pch = 1,
  ind.names = FALSE,
  legend = FALSE,
  background = infx_train_background,
  col = c("goldenrod", "gray75"),
  star = TRUE,
  point.lwd = 0.5,
  title = NULL,
  size.title = 0.00001,
  style = "graphics",
  legend.title = "Infection Group",
  X.label = paste0("Component 1 (", round(infx_train_splsda_final$prop_expl_var$X[1] * 100), "%)"),
  Y.label = paste0("Component 2 (", round(infx_train_splsda_final$prop_expl_var$X[2] * 100), "%)")
) 
# addtable2plot(
#   y = max(infx_train_splsda_final$variates$X[,2])*0.1,
#   x = min(infx_train_splsda_final$variates$X[,1]),
#   infx_confusion_df[,1:2],
#   bty = "o",
#   display.rownames = TRUE,
#   hlines = TRUE,
#   vlines = TRUE,
#   cex = 1.35,
#   bg = "white"
# )
text(
  y = -4,
  x = -2.75,
  substitute(bold("Infection")), cex = 1.75, adj = 0, col = "goldenrod")
text(
  y = 4,
  x = -1,
  substitute(bold("No Infection")), cex = 1.75, adj = 0, col = "gray70")
text(
  y = max(infx_train_splsda_final$variates$X[,2])*1,
  x = min(infx_train_splsda_final$variates$X[,1]),
  paste0("ACC = ", 
                    round(infx_cm$overall[1]*100, 1), "% [",
                    round(infx_cm$overall[3]*100, 1), "%, ",
                    round(infx_cm$overall[4]*100, 1), "%]"),
    cex = 1.35, adj = 0)
text(
  y = max(infx_train_splsda_final$variates$X[,2])*0.9,
  x = min(infx_train_splsda_final$variates$X[,1]),
  paste0("Sens. = ", 
                    paste0(formatC(round(infx_epi$detail[2][3,], 3)*100, digits = 1, format = "f"), "%"), " [",
                    paste0(formatC(round(infx_epi$detail[3][3,], 3)*100, digits = 1, format = "f"), "%"),", ",
                    paste0(formatC(round(infx_epi$detail[4][3,], 3)*100, digits = 1, format = "f"), "%"),"]"),
    cex = 1.35, adj = 0)
text(
  y = max(infx_train_splsda_final$variates$X[,2])*0.8,
  x = min(infx_train_splsda_final$variates$X[,1]),
  paste0("Spec. = ", 
                    paste0(formatC(round(infx_epi$detail[2][4,], 3)*100, digits = 1, format = "f"), "%"), " [",
                    paste0(formatC(round(infx_epi$detail[3][4,], 3)*100, digits = 1, format = "f"), "%"),", ",
                    paste0(formatC(round(infx_epi$detail[4][4,], 3)*100, digits = 1, format = "f"), "%"),"]"),
    cex = 1.35, adj = 0)
text(
  y = max(infx_train_splsda_final$variates$X[,2])*0.7,
  x = min(infx_train_splsda_final$variates$X[,1]),
  paste0("OR = ",
                    formatC(round(infx_epi$detail[2][6,], 3), digits = 1, format = "f"),  " [",
                    formatC(round(infx_epi$detail[3][6,], 3), digits = 1, format = "f"), ", ",
                    formatC(round(infx_epi$detail[4][6,], 3), digits = 1, format = "f"), "]"),
    cex = 1.35, adj = 0)

invisible(dev.off())
}


# Figure 6A
par(mfrow=c(1,2))

plotIndiv(
  diversity_train_splsda_final,
  xlim = c(min(diversity_train_splsda_final$variates$X[,1])*1.05, max(diversity_train_splsda_final$variates$X[,1])*1.8),
  ylim = c(min(diversity_train_splsda_final$variates$X[,2])*1.15, max(diversity_train_splsda_final$variates$X[,2])*1.8),
  comp = c(1,2),
  pch = 1,
  ind.names = FALSE,
  legend = FALSE,
  background = diversity_train_background,
  col = c("#3A001E", "#8A0246", "#C20463"),
  star = TRUE,
  point.lwd = 0.5,
  title = NULL,
  size.title = 0.00001,
  style = "graphics",
  legend.title = "Diversity Group",
  X.label = paste0("Component 1 (", round(diversity_train_splsda_final$prop_expl_var$X[1] * 100), "%)"),
  Y.label = paste0("Component 2 (", round(diversity_train_splsda_final$prop_expl_var$X[2] * 100), "%)")
) 
addtable2plot(
  y = -4.68,
  x = min(diversity_train_splsda_final$variates$X[,1])*1.1,
  diversity_confusion_df,
  bty = "o",
  display.rownames = TRUE,
  hlines = TRUE,
  vlines = TRUE,
  cex = 0.75,
  bg = "white"
)
text(
  y = -6.5,
  x = min(diversity_train_splsda_final$variates$X[,1])*1.1,
  paste0("Overall ACC = ", 
                    paste0(formatC(round(diversity_cm$overall[1], 3)*100, digits = 1, format = "f"), "%"), " [",
                    paste0(formatC(round(diversity_cm$overall[3], 3)*100, digits = 1, format = "f"), "%"),", ",
                    paste0(formatC(round(diversity_cm$overall[4], 3)*100, digits = 1, format = "f"), "%"),"]"),
    cex = 1.05, adj = 0)

text(
  y = 3,
  x = 0.5,
  paste0("Low Diversity ACC = ", round(diversity_epi$balanced.accuracy[1]*100, 1), "%"),
    cex = 1.05, adj = 0, col = "#3A001E")
text(
  y = 2.7,
  x = 0.5,
  paste0("Low Diversity Sens. = ", round(diversity_epi$precision[1]*100, 1), "%"),
    cex = 1.05, adj = 0, col = "#3A001E")
text(
  y = 2.4,
  x = 0.5,
  paste0("Low Diversity Spec. = ", round(diversity_epi$specificity[1]*100, 1), "%"),
    cex = 1.05, adj = 0, col = "#3A001E")
# text(
#   y = 2.1,
#   x = 0.5,
#   paste0("Low Diversity OR = ", round(diversity_epi$DOR[1], 1)),
#     cex = 1.05, adj = 0, col = "#3A001E")

text(
  y = -5.5,
  x = 0,
  paste0("Medium Diversity ACC = ", round(diversity_epi$balanced.accuracy[2]*100, 1), "%"),
    cex = 1.05, adj = 0, col = "#8A0246")
text(
  y = -5.8,
  x = 0,
  paste0("Medium Diversity Sens. = ", round(diversity_epi$precision[2]*100, 1), "%"),
    cex = 1.05, adj = 0, col = "#8A0246")
text(
  y = -6.1,
  x = 0,
  paste0("Medium Diversity Spec. = ", round(diversity_epi$specificity[2]*100, 1), "%"),
    cex = 1.05, adj = 0, col = "#8A0246")
# text(
#   y = -7.4,
#   x = 0,
#   paste0("Medium Diversity OR = ", round(diversity_epi$DOR[2], 1)),
#     cex = 1.05, adj = 0, col = "#8A0246")

text(
  y = 3,
  x = min(diversity_train_splsda_final$variates$X[,1])*1.05,
  paste0("High Diversity ACC = ", round(diversity_epi$balanced.accuracy[3]*100, 1), "%"),
    cex = 1.05, adj = 0, col = "#C20463")
text(
  y = 2.7,
  x = min(diversity_train_splsda_final$variates$X[,1])*1.05,
  paste0("High Diversity Sens. = ", round(diversity_epi$precision[3]*100, 1), "%"),
    cex = 1.05, adj = 0, col = "#C20463")
text(
  y = 2.4,
  x = min(diversity_train_splsda_final$variates$X[,1])*1.05,
  paste0("High Diversity Spec. = ", round(diversity_epi$specificity[3]*100, 1), "%"),
    cex = 1.05, adj = 0, col = "#C20463")
# text(
#   y = 2.1,
#   x = min(diversity_train_splsda_final$variates$X[,1])*1.05,
#   paste0("High Diversity OR = ", round(diversity_epi$DOR[3], 1)),
#     cex = 1.05, adj = 0, col = "#C20463")

# Figure 6B
plotIndiv(
  infx_train_splsda_final,
  comp = c(1,2),
  pch = 1,
  ind.names = FALSE,
  legend = FALSE,
  background = infx_train_background,
  col = c("goldenrod", "gray75"),
  star = TRUE,
  point.lwd = 0.5,
  title = NULL,
  size.title = 0.00001,
  style = "graphics",
  legend.title = "Infection Group",
  X.label = paste0("Component 1 (", round(infx_train_splsda_final$prop_expl_var$X[1] * 100), "%)"),
  Y.label = paste0("Component 2 (", round(infx_train_splsda_final$prop_expl_var$X[2] * 100), "%)")
) 
addtable2plot(
  y = max(infx_train_splsda_final$variates$X[,2])*0.1,
  x = min(infx_train_splsda_final$variates$X[,1]),
  infx_confusion_df[,1:2],
  bty = "o",
  display.rownames = TRUE,
  hlines = TRUE,
  vlines = TRUE,
  cex = 0.75,
  bg = "white"
)
text(
  y = -4,
  x = -2.75,
  substitute(bold("Infection")), cex = 1.75, adj = 0, col = "goldenrod")
text(
  y = 4,
  x = -1,
  substitute(bold("No Infection")), cex = 1.75, adj = 0, col = "gray70")
text(
  y = max(infx_train_splsda_final$variates$X[,2])*1,
  x = min(infx_train_splsda_final$variates$X[,1]),
  paste0("ACC = ", 
                    round(infx_cm$overall[1]*100, 1), "% [",
                    round(infx_cm$overall[3]*100, 1), "%, ",
                    round(infx_cm$overall[4]*100, 1), "%]"),
    cex = 1.05, adj = 0)
text(
  y = max(infx_train_splsda_final$variates$X[,2])*0.9,
  x = min(infx_train_splsda_final$variates$X[,1]),
  paste0("Sens. = ", 
                    paste0(formatC(round(infx_epi$detail[2][3,], 3)*100, digits = 1, format = "f"), "%"), " [",
                    paste0(formatC(round(infx_epi$detail[3][3,], 3)*100, digits = 1, format = "f"), "%"),", ",
                    paste0(formatC(round(infx_epi$detail[4][3,], 3)*100, digits = 1, format = "f"), "%"),"]"),
    cex = 1.05, adj = 0)
text(
  y = max(infx_train_splsda_final$variates$X[,2])*0.8,
  x = min(infx_train_splsda_final$variates$X[,1]),
  paste0("Spec. = ", 
                    paste0(formatC(round(infx_epi$detail[2][4,], 3)*100, digits = 1, format = "f"), "%"), " [",
                    paste0(formatC(round(infx_epi$detail[3][4,], 3)*100, digits = 1, format = "f"), "%"),", ",
                    paste0(formatC(round(infx_epi$detail[4][4,], 3)*100, digits = 1, format = "f"), "%"),"]"),
    cex = 1.05, adj = 0)
text(
  y = max(infx_train_splsda_final$variates$X[,2])*0.7,
  x = min(infx_train_splsda_final$variates$X[,1]),
  paste0("OR = ",
                    formatC(round(infx_epi$detail[2][6,], 3), digits = 1, format = "f"),  " [",
                    formatC(round(infx_epi$detail[3][6,], 3), digits = 1, format = "f"), ", ",
                    formatC(round(infx_epi$detail[4][6,], 3), digits = 1, format = "f"), "]"),
    cex = 1.05, adj = 0)
```

### Supplemental Figure 3A + 3B: Plot Loadings for Diversity Groups
```{r, supplemental-figure-3-splsda-diversity-groups, fig.dim = c(14, 10), message=FALSE, warning=FALSE}
# Take absolute value to make diversity group loadings plot narrower
diversity_train_splsda_final$loadings$X <- abs(diversity_train_splsda_final$loadings$X)

# Combine Supplemental Figures 3A + 3B

{pdf(file = "./Results/Supplemental_Figure_3.pdf",
         height = 8, 
         width = 11)

par(mfrow = c(1,3))

# Supplmental Figure 3A
plotLoadings(
  diversity_train_splsda_final,
  contrib = 'max',
  method = 'mean',
  comp = 1,
  legend = FALSE,
  legend.col = c("#EDE342", "#F69A97", "#FF51EB"),
  size.name = 1.1,
  size.title = rel(1),
  ndisplay = 50
)

# Supplmental Figure 3B
plotLoadings(
  diversity_train_splsda_final,
  contrib = 'max',
  method = 'mean',
  comp = 2,
  legend.col = c("#EDE342", "#F69A97", "#FF51EB"),
  size.name = 1.1,
  size.title = rel(1),
  ndisplay = 50
)

invisible(dev.off())
}

par(mfrow = c(1,3))

# Supplmental Figure 3A
plotLoadings(
  diversity_train_splsda_final,
  contrib = 'max',
  method = 'mean',
  comp = 1,
  legend = FALSE,
  legend.col = c("#EDE342", "#F69A97", "#FF51EB"),
  size.name = 0.6,
  size.title = rel(1),
  ndisplay = 50
)

# Supplmental Figure 3B
plotLoadings(
  diversity_train_splsda_final,
  contrib = 'max',
  method = 'mean',
  comp = 2,
  legend.col = c("#EDE342", "#F69A97", "#FF51EB"),
  size.name = 0.6,
  size.title = rel(1),
  ndisplay = 50
)
```

### Supplemental Figure 4A + 4B: Plot Loadings for Any Bacterial Infection
```{r, supplemental-figure-4-splsda-any-infection, fig.dim = c(14, 10), message=FALSE, warning=FALSE}
# Take absolute value to make diversity group loadings plot narrower
infx_train_splsda_final$loadings$X <- abs(infx_train_splsda_final$loadings$X)

# Combine Supplemental Figures 4A + 4B

{pdf(file = "./Results/Supplemental_Figure_4.pdf",
         height = 8, 
         width = 11)

par(mfrow = c(1,3))

# Supplmental Figure 4A
plotLoadings(
  infx_train_splsda_final,
  contrib = 'max',
  method = 'mean',
  comp = 1,
  legend = FALSE,
  legend.col = c("goldenrod", "gray75"),
  size.name = 1.1,
  size.title = rel(1),
  ndisplay = 50
)

# Supplmental Figure 4B
plotLoadings(
  infx_train_splsda_final,
  contrib = 'max',
  method = 'mean',
  comp = 2,
  legend.col = c("goldenrod", "gray75"),
  size.name = 1.1,
  size.title = rel(1),
  ndisplay = 50
)

invisible(dev.off())}

par(mfrow = c(1,3))

# Supplmental Figure 4A
plotLoadings(
  infx_train_splsda_final,
  contrib = 'max',
  method = 'mean',
  comp = 1,
  legend = FALSE,
  legend.col = c("goldenrod", "gray75"),
  size.name = 0.6,
  size.title = rel(1),
  ndisplay = 50
)

# Supplmental Figure 4B
plotLoadings(
  infx_train_splsda_final,
  contrib = 'max',
  method = 'mean',
  comp = 2,
  legend.col = c("goldenrod", "gray75"),
  size.name = 0.6,
  size.title = rel(1),
  ndisplay = 50
)

```

## Clinical Tables
### Supplemental Figure 2: Flow Diagram
```{r, supplemental-figure-2-flow-diagram, warning=FALSE, message=FALSE, fig.dim = c(12, 8)}
flow_chart <- 
  flow_exclusions(
  incl_counts = c(158, 130, 107, 25),
  total_label = "Total Patients Enrolled",
  incl_labels = c("Patients w/ Transplant", "Patients Included","Patients w/ Bacterial Infection"),
  excl_labels = c("No Transplant", "No Sample In Study Period\nDay -7 to +30", "Patients w/o Bacterial Infection"),
  show_count = TRUE
)

flow_chart

flow_chart %>% 
  export_svg() %>% 
  charToRaw() %>% 
  rsvg_pdf("./Results/Supplemental_Figure_2.pdf")
```

### Demographics
```{r, demographics, warning=FALSE, message=FALSE}
## Vector of variables to summarize
demo_vars <- c("race", "sex", "age", "meld_transplant",
               "Alcoholic Hepatitis", "Alcoholic Cirrhosis", "NAFLD/NASH", "Primary Sclerosing Cholangitis",
               "Acute Viral Hepatitis", "Chronic Hepatitis B", "Chronic Hepatitis C",
               "Autoimmune", "Wilson's Disease", 'Alpha-1 Antitrypsin', "Hemachromatosis", 
               "Drug Induced Liver Injury or Toxin", "Budd Chiari", "Cryptogenic", "Malignancy", "Other",  "Dialysis",
               "Pressers", "Mechanical Ventilation")
## Vector of categorical variables that need transformation
demo_cats <- c("race", "sex",
               "Alcoholic Hepatitis", "Alcoholic Cirrhosis", "NAFLD/NASH", "Primary Sclerosing Cholangitis",
               "Acute Viral Hepatitis", "Chronic Hepatitis B", "Chronic Hepatitis C",
               "Autoimmune", "Wilson's Disease", 'Alpha-1 Antitrypsin', "Hemachromatosis", 
               "Drug Induced Liver Injury or Toxin", "Budd Chiari", "Cryptogenic", "Malignancy", "Other",  "Dialysis",
               "Pressers", "Mechanical Ventilation")

  
tab1_1 <- CreateTableOne(
  vars = demo_vars,
  testNonNormal = "wilcox.test",
  includeNA = TRUE,
  factorVars = demo_cats,
  strata = "any_infection",
  data = demo
)

summary(tab1_1) # Age is potentially skewed, need to state that it is skewed and re-run `CreateTableOne`
tableone_skewed <- c("age", "meld_transplant")

tab1_2 <- print(tab1_1, nonnormal = tableone_skewed, formatOptions = list(big.mark = ","))

write.csv(tab1_2, "./Results/Demo_Table_1.csv", row.names = TRUE) # Saving then reading in the same data allows for an easy way to adjust p-values, since it loads the object as a dataframe

# Need to adjust pvalues and arrange properly....hence the multiple dataframes below
tab1_2_padjust1 <- read.csv("./Results/Demo_Table_1.csv") %>% 
  dplyr::rename(" " = X, "No Infection" = X0, "Bacterial Infection" = X1)

tab1_2_padjust2 <- tab1_2_padjust1 %>% 
  mutate(` ` = factor(` `, levels = tab1_2_padjust1$` `))

tab1_2_padjust3 <- tab1_2_padjust1 %>%
    mutate(test = ifelse(!is.na(p) & test == "", "chi.sq", test)) %>%
    group_by(test) %>%
    rstatix::adjust_pvalue(p.col = "p", method = "BH") %>% 
    ungroup() %>% 
    mutate(` ` = factor(` `, tab1_2_padjust2$` `)) %>% 
    arrange(` `) %>% 
    mutate(p = ifelse(is.na(p), "", p),
           p.adj = ifelse(is.na(p.adj), "", p.adj))

# Read in csv to then append adjusted pvalues
write.csv(tab1_2_padjust3,"./Results/Demo_Table_1_padjust.csv", row.names = FALSE)
```

### Culture: Supplemental Table 1
```{r, culture, warning=FALSE, message=FALSE}
# Percentage of infections # 
  ##  where there is an infection within 30 days after transplant and not 0 days before
cult_percent_infx_all <- peri_criteria_all %>% 
  filter(bact_infection_present == "Yes",
         between(eday, 0, 30)) %>% 
  group_by(patientID, sampleID, eday) %>% 
  arrange(-infx_stool) %>%  
  dplyr::slice(1) %>%
  ungroup() %>% 
  select(patientID, sampleID, bact_infection_present, infx_stool, organism1, micro1.factor) %>%
  distinct() %>% 
  mutate(organism1 = gsub(x = organism1, pattern = "\\s+", replacement = ""),
         organism1 = str_to_lower(string = organism1),
         organism1 = ifelse(grepl(x = organism1, pattern = "enterococcus|enterobacterales|klebsiella|escherichia|citrobacter|proteus|staphyl|clostrid|pseudo|steno|bacteroides|helico"), organism1, "Culture Negative")) %>% 
  group_by(patientID, sampleID, infx_stool) %>%
  mutate(`Enterococcus faecium` = grepl(x = organism1, pattern = "enterococcusfaecium", ignore.case = T),
         `Enterococcus faecalis` = grepl(x = organism1, pattern = "enterococcusfaecalis", ignore.case = T),
         `Enterococcus avium` = grepl(x = organism1, pattern = "enterococcusavium", ignore.case = T),
         `Enterococcus gallinarum` = grepl(x = organism1, pattern = "enterococcusgallinarum", ignore.case = T),
         `Klebsiella pneumoniae` = grepl(x = organism1, pattern = "klebsiellapneumoniae", ignore.case = T),
         `Enterobacter cloaceae` = grepl(x = organism1, pattern = "enterobactercloaceae", ignore.case = T),
         `Escherichia coli` = grepl(x = organism1, pattern = "escherichiacoli", ignore.case = T),
         `Citrobacter freundii` = grepl(x = organism1, pattern = "citrobacterfreundii", ignore.case = T),
         `Proteus mirabilis` = grepl(x = organism1, pattern = "proteusmirabilis", ignore.case = T),
         `Staphylococcus aureus` = grepl(x = organism1, pattern = "staphylococcusaureus", ignore.case = T),
         `Staphylococcus epidermis` = grepl(x = organism1, pattern = "staphylococcusepidermidis", ignore.case = T),
         `Pseudomonas aeruginosa` = grepl(x = organism1, pattern = "pseudomonasaeruginosa", ignore.case = T),
         `Stenotrophmonas maltophilia` = grepl(x = organism1, pattern = "stenotrophmonasmaltophilia", ignore.case = T),
         `Helicobacter pylori` = grepl(x = organism1, pattern = "helicobacterpylori", ignore.case = T),
         `Clostridium difficile` = grepl(x = organism1, pattern = "clostridiumdifficile|clostridioidesdifficile", ignore.case = T),
         `Bacteroides sp.` = grepl(x = organism1, pattern = "bacteroides", ignore.case = T),
         `Culture Negative` = grepl(x = organism1, pattern = "Culture Negative", ignore.case = T)) %>% 
  pivot_longer(-c(patientID:micro1.factor), names_to = "organisms", values_to = "org_presence") %>%
  mutate(organisms = ifelse(bact_infection_present == "No", "No Bacterial Infection", organisms),
         org_presence = ifelse(org_presence == TRUE, 1, 0)) %>% 
  group_by(sampleID, infx_stool, bact_infection_present, organisms) %>% 
  dplyr::slice_max(org_presence) %>% 
  ungroup() %>% 
  filter(organisms != "No Bacterial Infection") %>% 
  group_by(patientID, bact_infection_present, organisms) %>% 
  dplyr::slice_max(org_presence) %>% 
  ungroup() %>% 
  filter(org_presence == 1) %>%
  group_by(patientID, infx_stool, organisms) %>% 
  dplyr::slice(1) %>% 
  group_by(organisms) %>% 
  tally() %>% 
  ungroup() %>% 
  mutate(total_infections = sum(n)) %>% 
  replace_na(list(group="unknown")) %>%
  group_by(organisms) %>% 
  dplyr::summarize(percent = sum(n)/total_infections*100,
                   count = sum(n))
  
# Location of infections #
  ##  where there is an infection within 30 days after transplant and not 0 days before
cult_location_infx_all <- peri_criteria_all %>% 
  filter(bact_infection_present == "Yes",
         between(eday, 0, 30)) %>% 
  group_by(patientID, eday, key) %>% 
  arrange(-infx_stool) %>%  
  dplyr::slice(1) %>%
  ungroup() %>% 
  select(patientID, bact_infection_present, infx_stool, organism1, key) %>% 
  distinct() %>% 
  mutate(organism1 = gsub(x = organism1, pattern = "\\s+", replacement = ""),
         organism1 = str_to_lower(string = organism1),
         organism1 = ifelse(grepl(x = organism1, pattern = "enterococcus|enterobacterales|klebsiella|escherichia|citrobacter|proteus|staphyl|clostrid|pseudo|steno|bacteroides|helico"), organism1, "Culture Negative")) %>% 
  group_by(patientID, infx_stool) %>% 
  mutate(`Enterococcus faecium` = grepl(x = organism1, pattern = "enterococcusfaecium", ignore.case = T),
         `Enterococcus faecalis` = grepl(x = organism1, pattern = "enterococcusfaecalis", ignore.case = T),
         `Enterococcus avium` = grepl(x = organism1, pattern = "enterococcusavium", ignore.case = T),
         `Enterococcus gallinarum` = grepl(x = organism1, pattern = "enterococcusgallinarum", ignore.case = T),
         `Klebsiella pneumoniae` = grepl(x = organism1, pattern = "klebsiellapneumoniae", ignore.case = T),
         `Enterobacter cloaceae` = grepl(x = organism1, pattern = "enterobactercloaceae", ignore.case = T),
         `Escherichia coli` = grepl(x = organism1, pattern = "escherichiacoli", ignore.case = T),
         `Citrobacter freundii` = grepl(x = organism1, pattern = "citrobacterfreundii", ignore.case = T),
         `Proteus mirabilis` = grepl(x = organism1, pattern = "proteusmirabilis", ignore.case = T),
         `Staphylococcus aureus` = grepl(x = organism1, pattern = "staphylococcusaureus", ignore.case = T),
         `Staphylococcus epidermis` = grepl(x = organism1, pattern = "staphylococcusepidermidis", ignore.case = T),
         `Pseudomonas aeruginosa` = grepl(x = organism1, pattern = "pseudomonasaeruginosa", ignore.case = T),
         `Stenotrophmonas maltophilia` = grepl(x = organism1, pattern = "stenotrophmonasmaltophilia", ignore.case = T),
         `Helicobacter pylori` = grepl(x = organism1, pattern = "helicobacterpylori", ignore.case = T),
         `Clostridium difficile` = grepl(x = organism1, pattern = "clostridiumdifficile|clostridioidesdifficile", ignore.case = T),
         `Bacteroides sp.` = grepl(x = organism1, pattern = "bacteroides", ignore.case = T),
         `Culture Negative` = grepl(x = organism1, pattern = "Culture Negative", ignore.case = T)) %>% 
  pivot_longer(-c(patientID:key), names_to = "organisms", values_to = "org_presence") %>% 
  mutate(organisms = ifelse(bact_infection_present == "No", "No Bacterial Infection", organisms),
         org_presence = ifelse(org_presence == TRUE, 1, 0)) %>% 
 group_by(patientID, key, infx_stool, bact_infection_present) %>% 
  dplyr::slice_max(org_presence) %>% 
  ungroup() %>% 
  filter(org_presence == 1) %>% 
  filter(organisms != "No Bacterial Infection") %>%
  group_by(patientID, infx_stool, key) %>% 
  dplyr::slice(1) %>% 
  group_by(key) %>% 
  tally() %>% 
  ungroup() %>% 
  mutate(total_infections = sum(n)) %>% 
  replace_na(list(group="unknown")) %>%
  group_by(key) %>% 
  dplyr::summarize(percent = sum(n)/total_infections*100,
                   count = sum(n))

# Type of Infection #
  ##  where there is an infection within 30 days after transplant and not 0 days before
cult_type_infx_all <- peri_criteria_all %>% 
  filter(bact_infection_present == "Yes",
         grepl(variable, pattern = "anatomy.+"),
         between(eday, 0, 30)) %>% 
  group_by(patientID, eday, diag_cat2) %>% 
  arrange(-infx_stool) %>%  
  dplyr::slice(1) %>% 
  ungroup() %>% 
  mutate(diag_cat3 = case_when(grepl("abdominal", diag_cat2, ignore.case = T) ~ "Abdominal",
                               grepl("vir|bronchitis|covid-19|cmv", diag_cat2, ignore.case = T) ~ "Viral",
                               grepl("bacteremia", diag_cat2, ignore.case = T) ~ "Bacteremia",
                               grepl("thrush", diag_cat2, ignore.case = T) ~ "Fungal",
                               grepl("cholangitis|empyema|panniculitis|peritonitis|latent tb", diag_cat2, ignore.case = T) ~ "Bacterial",
                               grepl("cystitis|pyelonephritis", diag_cat2, ignore.case = T) ~ "Urinary Tract Infection",
                               grepl("pneumonia", diag_cat2, ignore.case = T) ~ "Pneumonia",
                               TRUE~diag_cat2),
         diag_cat3 = str_to_title(diag_cat3)) %>% 
  ungroup() %>% 
 select(patientID, bact_infection_present, infx_stool, organism1, micro1.factor, key, diag_cat3) %>% 
  distinct() %>% 
  mutate(organism1 = gsub(x = organism1, pattern = "\\s+", replacement = ""),
         organism1 = str_to_lower(string = organism1),
         organism1 = ifelse(grepl(x = organism1, pattern = "enterococcus|enterobacterales|klebsiella|escherichia|citrobacter|proteus|staphyl|clostrid|pseudo|steno|bacteroides|helico"), organism1, "Culture Negative")) %>% 
  group_by(patientID, infx_stool) %>% 
  mutate(`Enterococcus faecium` = grepl(x = organism1, pattern = "enterococcusfaecium", ignore.case = T),
         `Enterococcus faecalis` = grepl(x = organism1, pattern = "enterococcusfaecalis", ignore.case = T),
         `Enterococcus avium` = grepl(x = organism1, pattern = "enterococcusavium", ignore.case = T),
         `Enterococcus gallinarum` = grepl(x = organism1, pattern = "enterococcusgallinarum", ignore.case = T),
         `Klebsiella pneumoniae` = grepl(x = organism1, pattern = "klebsiellapneumoniae", ignore.case = T),
         `Enterobacter cloaceae` = grepl(x = organism1, pattern = "enterobactercloaceae", ignore.case = T),
         `Escherichia coli` = grepl(x = organism1, pattern = "escherichiacoli", ignore.case = T),
         `Citrobacter freundii` = grepl(x = organism1, pattern = "citrobacterfreundii", ignore.case = T),
         `Proteus mirabilis` = grepl(x = organism1, pattern = "proteusmirabilis", ignore.case = T),
         `Staphylococcus aureus` = grepl(x = organism1, pattern = "staphylococcusaureus", ignore.case = T),
         `Staphylococcus epidermis` = grepl(x = organism1, pattern = "staphylococcusepidermidis", ignore.case = T),
         `Pseudomonas aeruginosa` = grepl(x = organism1, pattern = "pseudomonasaeruginosa", ignore.case = T),
         `Stenotrophmonas maltophilia` = grepl(x = organism1, pattern = "stenotrophmonasmaltophilia", ignore.case = T),
         `Helicobacter pylori` = grepl(x = organism1, pattern = "helicobacterpylori", ignore.case = T),
         `Clostridium difficile` = grepl(x = organism1, pattern = "clostridiumdifficile|clostridioidesdifficile", ignore.case = T),
         `Bacteroides sp.` = grepl(x = organism1, pattern = "bacteroides", ignore.case = T),
         `Culture Negative` = grepl(x = organism1, pattern = "Culture Negative", ignore.case = T)) %>% 
  pivot_longer(-c(patientID:diag_cat3), names_to = "organisms", values_to = "org_presence") %>% 
  mutate(organisms = ifelse(bact_infection_present == "No", "No Bacterial Infection", organisms),
         org_presence = ifelse(org_presence == TRUE, 1, 0)) %>% 
  group_by(patientID, diag_cat3, infx_stool, bact_infection_present) %>% 
  dplyr::slice_max(org_presence) %>% 
  ungroup() %>% 
  filter(org_presence == 1,
         organisms != "No Bacterial Infection") %>%
  ungroup() %>% 
  group_by(patientID, infx_stool, diag_cat3) %>% 
  dplyr::slice(1) %>%
  group_by(diag_cat3) %>% 
  tally() %>% 
  ungroup() %>% 
  mutate(total_infections = sum(n)) %>% 
  group_by(diag_cat3) %>% 
  dplyr::summarize(percent = sum(n)/total_infections*100,
                   count = sum(n))

culture_tot <- bind_rows(cult_percent_infx_all,
                     cult_location_infx_all,
                      cult_type_infx_all) %>% 
  mutate(percent = round(percent, 2))

# Read in csv to then append adjusted pvalues
write.csv(culture_tot, "./Results/Supplemental_Table_1.csv", row.names = FALSE)
```

### Antibiotics
```{r, abx, warning=FALSE, message=FALSE}
## Vector of variables to summarize
abx_vars <- c("Basiliximab", "Mycophenolate", "Steroid", "Systemic Vancomycin", "Tacrolimus", "Cefepime", "Metronidazole", "Piperacillin/Tazobactam", "Rifaximin", "Ceftriaxone", "Ciprofloxacin", "Gentamicin", "Tobramicin",
              "Daptomycin", "Meropenem", "Oral Vancomycin")

abx2 <- abx %>% 
  filter(grepl(pattern = "basilix|tacro|steroid|mycophenolate|lactulose", medication_name, ignore.case = T)|
           grepl(pattern = "GLUCOCORTICOIDS|steroid", med_pharm_class, ignore.case = T)|
           grepl(pattern = "steroid", med_pharm_sub_class, ignore.case = T)|
           grepl("given", mar_action, ignore.case = T)) %>%
  mutate(Immunosuppressants = case_when(grepl("basilix", medication_name, ignore.case = T) ~ "Basiliximab",
                              grepl("tacro", medication_name, ignore.case = T) ~ "Tacrolimus",
                              grepl("one|ide|solu-cortef", medication_name, ignore.case = T) ~ "Steroid",
                              grepl("mycophenolate", medication_name, ignore.case = T) ~ "Mycophenolate",
                              grepl("lactulose", medication_name, ignore.case = T) ~ "Lactulose"),
         Antibiotics = case_when(grepl("rifaximin", medication_name, ignore.case = T) ~ "Rifaximin",
                              grepl("lactulose", medication_name, ignore.case = T) ~ "Lactulose",
                              grepl("ceftriaxone", medication_name, ignore.case = T) ~ "Ceftriaxone",
                              grepl("piperacillin|tazobactam", medication_name, ignore.case = T) ~ "Piperacillin/Tazobactam",
                              grepl("cefepime", medication_name, ignore.case = T) ~ "Cefepime",
                              grepl("meropenem", medication_name, ignore.case = T) ~ "Meropenem",
                              grepl("gentamicin", medication_name, ignore.case = T) ~ "Gentamicin",
                              grepl("tobramycin", medication_name, ignore.case = T) ~ "Tobramicin",
                              grepl("vancomycin.+oral", medication_name, ignore.case = T) ~ "Oral Vancomycin",
                              grepl("vancomycin.+(IV|Intravenous)", medication_name, ignore.case = T) ~ "Systemic Vancomycin",
                              grepl("METRONIDAZOLE", medication_name, ignore.case = T) ~ "Metronidazole",
                              grepl("DAPTOMYCIN", medication_name, ignore.case = T) ~ "Daptomycin",
                              grepl("linezolid", medication_name, ignore.case = T) ~ "Linezolid",
                              grepl("fluconazole", medication_name, ignore.case = T) ~ "Fluconazole",
                              grepl("micafungin", medication_name, ignore.case = T) ~ "Micafungin",
                              grepl("cipro", medication_name, ignore.case = T) & !grepl("drop", dose_units, ignore.case = T) ~ "Ciprofloxacin"),
         action = case_when(!is.na(Immunosuppressants) & between(days_transplant, 0, 30)|!is.na(Immunosuppressants) & ordering_mode == "Outpatient" ~ "keep",
                            !is.na(Antibiotics) & between(days_transplant, -7, 1) ~ "keep",
                            TRUE ~ "remove")) %>% 
  group_by(patientID, Immunosuppressants, Antibiotics) %>% 
  arrange(days_transplant) %>% 
  filter(action == "keep") %>% 
  dplyr::slice(1) %>% 
  select(patientID, Immunosuppressants, Antibiotics) %>% 
  left_join(peri_matrix_all %>% select(patientID, any_infection)) %>% 
  pivot_longer(!c(patientID, any_infection), names_to = "variable", values_to = "value") %>% 
  drop_na(value) %>% 
  mutate(variable = 1) %>% 
  pivot_wider(c(patientID, any_infection), names_from = "value", values_from = "variable", values_fn = min) %>% 
  replace(is.na(.), 0)

abx_tab1_1 <- CreateTableOne(
  vars = abx_vars,
  testNonNormal = "wilcox.test",
  includeNA = FALSE,
  factorVars = abx_vars,
  strata = "any_infection",
  data = abx2
)

summary(abx_tab1_1)

abx_tab1_2 <- print(abx_tab1_1, formatOptions = list(big.mark = ","))

write.csv(abx_tab1_2, "./Results/ABX_Table_1.csv", row.names = TRUE) # Saving then reading in the same data allows for an easy way to adjust p-values, since it loads the object as a dataframe

# Need to adjust pvalues and arrange properly....hence the multiple dataframes below
abx_tab1_2_padjust1 <- read.csv("./Results/ABX_Table_1.csv") %>% 
  dplyr::rename(" " = X, "No Infection" = X0, "Bacterial Infection" = X1)

abx_tab1_2_padjust2 <- abx_tab1_2_padjust1 %>% 
  mutate(` ` = factor(` `, levels = abx_tab1_2_padjust1$` `))

abx_tab1_2_padjust3 <- abx_tab1_2_padjust1 %>%
    mutate(test = ifelse(!is.na(p) & is.na(test), "chi.sq", "")) %>%
    group_by(test) %>%
    rstatix::adjust_pvalue(p.col = "p", method = "BH") %>% 
    ungroup() %>% 
    mutate(` ` = factor(` `, abx_tab1_2_padjust2$` `)) %>% 
    arrange(` `) %>% 
    mutate(p = ifelse(is.na(p), "", p),
           p.adj = ifelse(is.na(p.adj), "", p.adj))

# Read in csv to then append adjusted pvalues
write.csv(abx_tab1_2_padjust3, "./Results/ABX_Table_1_padjust.csv", row.names = FALSE)
```

### Miscellaneous Statistics: Supplemental Table 2
```{r, miscellaneous-stats, warning=FALSE, message=FALSE}
# Expansions and Infections Stats
expan_infx_stats <-
peri_matrix_all %>% 
  left_join(peri_matrix_all %>% select(sampleID, ends_with("infection")) %>% ungroup()) %>% 
  mutate(ecoc_infx = enterococcus_infection,
         ecoc_infx = ifelse(ecoc_infx >= 1, 1, 0),
         ebac_infx = enterobacterales_infection,
         ebac_infx = ifelse(ebac_infx >= 1, 1, 0)) %>% 
  select(enterococcus_rel_abundance, enterobacterales_rel_abundance, ecoc_infx, ebac_infx) %>% 
  summarise(ecoc_expan_above_cutpoint = sum(enterococcus_rel_abundance >= optimal_cutpoint_rel$optimal_cutpoint[2]),
            ecoc_infx_ecoc_below_cutpoint = sum(ecoc_infx == 1 & enterococcus_rel_abundance < optimal_cutpoint_rel$optimal_cutpoint[2], na.rm=T),
            ecoc_infx_ecoc_above_cutpoint = sum(ecoc_infx == 1 & enterococcus_rel_abundance >= optimal_cutpoint_rel$optimal_cutpoint[2], na.rm=T),
            ecoc_noinfx_ecoc_below_cutpoint = sum(ecoc_infx == 0 & enterococcus_rel_abundance < optimal_cutpoint_rel$optimal_cutpoint[2], na.rm=T),
            ecoc_noinfx_ecoc_above_cutpoint = sum(ecoc_infx == 0 & enterococcus_rel_abundance >= optimal_cutpoint_rel$optimal_cutpoint[2], na.rm=T),
            ebac_expan_above_cutpoint = sum(enterobacterales_rel_abundance >= optimal_cutpoint_rel$optimal_cutpoint[1]),
            ebac_infx_ebac_below_cutpoint = sum(ebac_infx == 1 & enterobacterales_rel_abundance < optimal_cutpoint_rel$optimal_cutpoint[1], na.rm=T),
            ebac_infx_ebac_above_cutpoint = sum(ebac_infx == 1 & enterobacterales_rel_abundance >= optimal_cutpoint_rel$optimal_cutpoint[1], na.rm=T),
            ebac_noinfx_ebac_below_cutpoint = sum(ebac_infx == 0 & enterobacterales_rel_abundance < optimal_cutpoint_rel$optimal_cutpoint[1], na.rm=T),
            ebac_noinfx_ebac_above_cutpoint = sum(ebac_infx == 0 & enterobacterales_rel_abundance >= optimal_cutpoint_rel$optimal_cutpoint[1], na.rm=T)) %>%
  rownames_to_column(var = "rowname") %>% 
  pivot_longer(-rowname, names_to = "metric", values_to = "count") %>% 
  mutate(percent = round((count / 107) * 100, 2)) %>% 
  select(-rowname)

ecoc_infx_confusion <- expan_infx_stats %>% 
  filter(metric %in% c(
    "ecoc_infx_ecoc_below_cutpoint",
    "ecoc_infx_ecoc_above_cutpoint",
    "ecoc_noinfx_ecoc_below_cutpoint",
    "ecoc_noinfx_ecoc_above_cutpoint"
  )) %>% 
  select(metric, count) %>% 
  pivot_wider(names_from = metric, values_from = count)

ecoc_infx_confusion_cnfs <- data.frame("Enterococcus" = c("Infection", "No Infection"),
                                       "Greater Than Cutpoint" = c(ecoc_infx_confusion$ecoc_infx_ecoc_above_cutpoint,
                                                                   ecoc_infx_confusion$ecoc_noinfx_ecoc_above_cutpoint),
                                       "Less Than Cutpoint" = c(ecoc_infx_confusion$ecoc_infx_ecoc_below_cutpoint,
                                                                ecoc_infx_confusion$ecoc_noinfx_ecoc_below_cutpoint))

ecoc_infx_confusion_cnfs <- ecoc_infx_confusion_cnfs %>% 
  mutate(sensitivity = round(ecoc_infx_confusion_cnfs[1,2]/(ecoc_infx_confusion_cnfs[1,2] + ecoc_infx_confusion_cnfs[1,3]), 3),
         specificity = round(ecoc_infx_confusion_cnfs[2,3]/(ecoc_infx_confusion_cnfs[2,3] + ecoc_infx_confusion_cnfs[2,2]), 3),
         odds_ratio = round((ecoc_infx_confusion_cnfs[1,2]/ecoc_infx_confusion_cnfs[1,3])/(ecoc_infx_confusion_cnfs[2,2]/ecoc_infx_confusion_cnfs[2,3]), 3))

write.csv(ecoc_infx_confusion_cnfs, "./Results/Supplemental_Table_2A.csv", row.names = FALSE)


ebac_infx_confusion <- expan_infx_stats %>% 
  filter(metric %in% c(
    "ebac_infx_ebac_below_cutpoint",
    "ebac_infx_ebac_above_cutpoint",
    "ebac_noinfx_ebac_below_cutpoint",
    "ebac_noinfx_ebac_above_cutpoint"
  )) %>% 
  select(metric, count) %>% 
  pivot_wider(names_from = metric, values_from = count)

ebac_infx_confusion_cnfs <- data.frame("Enterobacterales" = c("Infection", "No Infection"),
                                       "Greater Than Cutpoint" = c(ebac_infx_confusion$ebac_infx_ebac_above_cutpoint,
                                                                   ebac_infx_confusion$ebac_noinfx_ebac_above_cutpoint),
                                       "Less Than Cutpoint" = c(ebac_infx_confusion$ebac_infx_ebac_below_cutpoint,
                                                                ebac_infx_confusion$ebac_noinfx_ebac_below_cutpoint))

ebac_infx_confusion_cnfs <- ebac_infx_confusion_cnfs %>% 
  mutate(sensitivity = round(ebac_infx_confusion_cnfs[1,2]/(ebac_infx_confusion_cnfs[1,2] + ebac_infx_confusion_cnfs[1,3]), 3),
         specificity = round(ebac_infx_confusion_cnfs[2,3]/(ebac_infx_confusion_cnfs[2,3] + ebac_infx_confusion_cnfs[2,2]), 3),
         odds_ratio = round((ebac_infx_confusion_cnfs[1,2]/ebac_infx_confusion_cnfs[1,3])/(ebac_infx_confusion_cnfs[2,2]/ebac_infx_confusion_cnfs[2,3]), 3))

write.csv(ebac_infx_confusion_cnfs, "./Results/Supplemental_Table_2B.csv", row.names = FALSE)
```

### Supplemental Figure 1: Fecal Sample Collected from Transplant
```{r, sample-relative-to-transplant,warning=FALSE, message=FALSE, fig.dim = c(12, 8)}
sample_days <- 
  metaphlan_df_sumry %>% 
  group_by(sampleID) %>% 
  slice(1) %>% 
  filter(db == "Liver Transplant") %>% 
  select(sampleID, diversity_group) %>% 
  left_join(sample_lookup %>% 
              select(sampleID, sample_days_from_transplant)) %>% 
  ungroup()


# Quartiles of samples
clin_quartiles <-
  sample_days %>% 
  select(sample_days_from_transplant) %>% 
  summarise(minimum = min(sample_days_from_transplant, na.rm = T),
            lower = unname(quantile(sample_days_from_transplant, probs = .25, na.rm = T)),
            median = unname(quantile(sample_days_from_transplant, probs = .5, na.rm = T)),
            upper = unname(quantile(sample_days_from_transplant, probs = 0.75, na.rm = T)),
            maximum = max(sample_days_from_transplant, na.rm = T)) %>% 
  mutate(placeholder = 1) %>% 
  pivot_longer(!placeholder, names_to = "ranges", values_to = "value")



sample_days_histogram <-
sample_days %>% 
  ggplot(., aes(x = sample_days_from_transplant)) +
  geom_histogram(color="black", fill="grey", binwidth = 1) +
  geom_vline(aes(xintercept = 0), color="orangered", linetype="solid", size = 1.2) +
  geom_vline(aes(xintercept = median(sample_days_from_transplant, na.rm = TRUE)), color = "cyan2", linetype="dashed", size = 1.2) +
  geom_text(aes(x = median(sample_days_from_transplant, na.rm = TRUE), y = 12.5), color = "cyan2", label = paste0("Median = ",
                                                                                           clin_quartiles %>%
                                                                                             filter(ranges == "median") %>%
                                                                                             select(value),
                                                                                           " days"), nudge_x = 4) +
  geom_text(label = "Transplant Date = Day 0", x = -4, y = 12.5, color="orangered") +
  theme_bw()+
  theme(
    panel.grid = eb(),
    axis.text = et(color = "black", size = 12),
    axis.title = et(color = "black", size = 14),
    legend.position = "none"
  ) +
  scale_x_continuous(breaks = seq(-7, 30, 5),
                     limits = c(-7, 35)) +
  ylab("Number of Samples\n")
 

sample_days_boxplot <-
sample_days %>% 
ggplot(., aes(x = sample_days_from_transplant, y = diversity_group, fill = diversity_group)) + 
  geom_vline(aes(xintercept = 0), color="orangered", linetype="solid", size = 1.2) +
  geom_vline(aes(xintercept = median(sample_days_from_transplant, na.rm = TRUE)), color="cyan2", linetype="dashed", size = 1.2) +
  geom_violin() +
  geom_boxplot(alpha = 0.5, width = 0.15, outlier.shape = NA, fill = "white", color = "white") +
  stat_compare_means(comparisons = list(c("High Diversity", "Medium Diversity"),
                                        c("High Diversity", "Low Diversity"),
                                        c("Medium Diversity", "Low Diversity")),
                      bracket.size = 0.3,
                     step.increase = 0.07
                     ) +
  theme_bw() +
  theme(
    panel.grid = eb(),
    axis.text.x = et(color = "black", size = 12),
    axis.text.y = et(color = "black", size = 12),
    axis.ticks.y = eb(),
    axis.title = et(color = "black", size = 14),
    legend.position = "none"
  ) +
  scale_x_continuous(breaks = seq(-7, 30, 5),
                     limits = c(-7, 35)) +
  scale_fill_manual(values = c("#3A001E", "#8A0246", "#C20463")) +
  ylab("Diversity Groups\n") +
  xlab("\nDays to Sample")

pdf(file = "./Results/Supplemental_Figure_1.pdf", height = 10, width = 12, onefile = F)

gg.stack(sample_days_histogram,
         sample_days_boxplot)

dev.off()

gg.stack(sample_days_histogram,
         sample_days_boxplot)
```

### Save Data Image
```{r save image, message=FALSE, warning=FALSE}
save.image(file = "./Data/LT_Modeling.RData")
```